Lesson 3
========================================================

***

### What to Do First?
Notes:

***

### Pseudo-Facebook User Data
Notes:

```{r Pseudo-Facebook User Data}

pf <-  read.csv(
  "pseudo_facebook.tsv",
  sep="\t")
names(pf)
str(pf)

```

***

### Histogram of Users' Birthdays
Notes:

```{r Histogram of Users\' Birthdays}
library(ggplot2)

ggplot(data = pf, aes(x = dob_day)) + 
  geom_histogram(binwidth = 1) + 
  scale_x_continuous(breaks = 1:31) + 
  facet_wrap(~dob_month)
```

***

#### What are some things that you notice about this histogram?
Response:
Outlier in January 1st
***

### Moira's Investigation
Notes:

***

### Estimating Your Audience Size
Notes:

***

#### Think about a time when you posted a specific message or shared a photo on Facebook. What was it?
Response:

#### How many of your friends do you think saw that post?
Response:

#### Think about what percent of your friends on Facebook see any posts or comments that you make in a month. What percent do you think that is?
Response:

***

### Perceived Audience Size
Notes:

***
### Faceting
Notes:

```{r Faceting}

```

#### Let’s take another look at our plot. What stands out to you here?
Response:

***

### Be Skeptical - Outliers and Anomalies
Notes:

***

### Moira's Outlier
Notes:
#### Which case do you think applies to Moira’s outlier?
Response:

***

### Friend Count
Notes:

#### What code would you enter to create a histogram of friend counts?

```{r Friend Count}

library(ggplot2)

ggplot(data = pf, mapping=aes(x = friend_count)) + 
  geom_histogram(binwidth = 1) +
  xlim(0, 1000)

```

```{r Friend Count2}



ggplot(data = pf, mapping = aes(x = friend_count)) +
geom_histogram(binwidth = 25) +
scale_x_continuous(limits = c(0, 1000), breaks = seq(0, 10000,50))

```

#### How is this plot similar to Moira's first plot?
Response:
positively scewed
***

### Limiting the Axes
Notes:

```{r Limiting the Axes}

```

### Exploring with Bin Width
Notes:

***

### Adjusting the Bin Width
Notes:

### Faceting Friend Count
```{r Faceting Friend Count}
# What code would you add to create a facet the histogram by gender?
# Add it to the code below.
qplot(x = friend_count,
      data = pf,
      binwidth = 10) +
      scale_x_continuous(limits = c(0, 1000),
      breaks = seq(0, 1000, 50)) +
      facet_wrap( ~ gender)
```

***

### Omitting NA Values
Notes:

```{r Omitting NA Values}

qplot(x = friend_count,
      data = subset(pf,!is.na(gender)),
      binwidth = 10) +
      scale_x_continuous(limits = c(0, 1000),
      breaks = seq(0, 1000, 50)) +
      facet_wrap( ~ gender)

```

***

### Statistics 'by' Gender
Notes:

```{r Statistics \'by\' Gender}
table(pf$gender)
by(pf$friend_count, pf$gender, summary)

```

#### Who on average has more friends: men or women?
Response:
women

#### What's the difference between the median friend count for women and men?
Response:
22
#### Why would the median be a better measure than the mean?
Response:
not so much influenced by outlier
***

### Tenure
Notes:

```{r Tenure}
ggplot(aes(x = tenure), data = pf) + 
   geom_histogram(binwidth = 30, color = 'black', fill = '#099DD9')
```

***

#### How would you create a histogram of tenure by year?

```{r Tenure Histogram by Year}
ggplot(aes(x = tenure / 365),
       data = pf) +
       geom_histogram(binwidth = .25,
       color = 'black',
       fill = '#F79420') +
       scale_x_continuous(limits = c(0, 7),
       breaks = seq(0, 7, .5)) +
       xlab("Number of years on Facebook") +
       ylab('Number of users')
  
```

***

### Labeling Plots
Notes:

```{r Labeling Plots}

ggplot(aes(x = tenure/365), data = pf) + 
   geom_histogram(binwidth = .25, color = 'black', fill = '#F79420')+
      scale_x_continuous(limits = c(0, 7),
      breaks = seq(0, 7, .5))

```

***

### User Ages
Notes:

```{r User Ages}

ggplot(data = pf, mapping = aes(x = age)) +
  geom_histogram(binwidth = 1) +
  scale_x_continuous(limits = c(0, 100),
  breaks = seq(0, 100, 2))


```{r User Ages}

summary(pf$age)
```
#### What do you notice?
Response:

***

### The Spread of Memes
Notes:

***

### Lada's Money Bag Meme
Notes:

***

### Transforming Data
Notes:

***

### Add a Scaling Layer
Notes:

```{r Add a Scaling Layer}




library(gridExtra)

p1 = ggplot(data = pf, mapping = aes(x = friend_count)) +
  geom_histogram(binwidth = 1,
                 color = 'black',
                 fill = '#099DD9')


p2 = p1 + scale_x_log10()


p3 = p1 + scale_x_sqrt()

grid.arrange(p1, p2, p3, ncol = 1)
```

***


### Frequency Polygons

```{r Frequency Polygons}

p1 = ggplot(aes(x = friend_count, y = ..count.. / sum(..count..)), data = subset(pf, !is.na(gender))) +
geom_freqpoly(aes(color = gender), binwidth = 10) +
xlab('Friend Count') +
ylab('Percentage of users with that friend count')

p2 = p1 +
scale_x_continuous(limits = c(0, 1000), breaks = seq(0, 1000, 50)) 


p3 = p1 + scale_x_log10()


grid.arrange(p2, p3, ncol = 1)
```

***

### Likes on the Web
Notes:

```{r Likes on the Web}

p1 = ggplot(aes(x = www_likes), data = subset(pf,!is.na(gender))) +
geom_freqpoly(aes(color = gender)) + scale_x_continuous()

p2 = p1 + scale_x_log10()



grid.arrange(p1, p2, ncol = 1)

```

```{r code}


by(pf$www_likes, pf$gender, sum)

```




***

### Box Plots
Notes:

```{r Box Plots}





ggplot(data = subset(pf, !is.na(gender)),
mapping = aes(x = gender, y = friend_count)) +
geom_boxplot() + scale_y_continuous(limits = c(0, 1000))



ggplot(data = subset(pf,!is.na(gender)),
mapping = aes(x = gender, y = friend_count)) +
geom_boxplot() + coord_cartesian(ylim = c(0, 250))
```

```{r}

by(pf$friend_count,pf$gender, summary)

```

#### Adjust the code to focus on users who have friend counts between 0 and 1000.

```{r}

```

***

### Box Plots, Quartiles, and Friendships
Notes:

```{r Box Plots, Quartiles, and Friendships}

```

#### On average, who initiated more friendships in our sample: men or women?
Response:
#### Write about some ways that you can verify your answer.
Response:
```{r Friend Requests by Gender}



ggplot(
data = subset(pf,!is.na(gender)),
mapping = aes(x = gender, y = friendships_initiated)
) +
geom_boxplot() + coord_cartesian(ylim = c(0, 125))

```

Response:

***

### Getting Logical
Notes:

```{r Getting Logical}



pf$mobile_check_in <- NA
pf$mobile_check_in <- ifelse(pf$mobile_likes > 0, 1, 0)
pf$mobile_check_in <- factor(pf$mobile_check_in)

summary(pf$mobile_check_in)

length(subset(pf$mobile_check_in,pf$mobile_check_in == 1)) / length(pf$mobile_check_in)

```

Response:

***

### Analyzing One Variable
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!
