---
title: Exploration of factors which influence the yearly revenue of an Airbnb listing
  on Bali by Jens Laufer
output:
  pdf_document: default
  html_notebook: default
  html_document: default
---

Everybody heard about Airbnb, which is the fastest growing holiday website in the world. 

I am Airbnb host since 2015 and indeed it was one of the factors that lead me into the fascinating world of data science as I want to improve my own listings and find new opportunities to invest.

Currently I am helping an agency on Bali, Indonesia to improve their listings  to gain experience with data analysis. 

I want to answer the question what factors influence the yearly revenue of an Airbnb listing on Bali.

The dataset for this project is scraped from Airbnb through their internal REST Api with python scripts, which are not part of this analysis. It includes 1236 listings on Bali and the booking calendar for the next 6 months with about 230,000 entries.




```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
library(ggplot2)
library(tidyr)
library(dplyr)
library(GGally)
library(gridExtra)
```

```{r echo=FALSE, message=FALSE, warning=FALSE, Data_Import}
listings_df <- read.csv('data/airbnb.listings.bali.csv')
calendar_df <- read.csv('data/airbnb.listing.calendar.dates.bali.csv')

```


```{r echo=FALSE, message=FALSE, warning=FALSE, Data_Modification}
# I am transforming the available into a Boolean value
calendar_df$available = as.logical(calendar_df$available == "true")

```





```{r echo=FALSE, warning=FALSE, message=FALSE, Data_Wrangling}
# Based on the "low level" calendar data I am calculating several metrics, 
# which are important for the hotel industry  (http://www.traveltripper.com/blog/the-most-important-metrics-for-hotel-revenue-managers/)
# These metrics are:
# Occupancy (OCC): Number of days a listing is not available / Total number of days
# Revenue: The total revenue in the period
# Average daily rate (ADR): Total revenue / Number of days a listing is not available
# Revenue per availabe room (RevPAR): Total revenue / Total number of days
# Yearly Revenue (estimated): 365 x Revenue per available room

 daily_revenue <- function(available, price) {
    ifelse(available == FALSE, price, 0)
 }

 average_daily_rate <- function(available, revenue) {
   if(length(which(available==FALSE)) > 0){
      revenue / length(which(available==FALSE))
   }
   else{
      0
   }
 }

 occupancy <- function(available) {
    length(which(available==FALSE)) / length(available)
 }
 
 revenue_per_available_room <- function(available, revenue) {
    average_daily_rate(available, revenue) * occupancy(available)
 }
 
 yearly_revenue <- function(available, revenue) {
    365 * revenue_per_available_room(available,revenue)
 }
 
# Calculation of metrics and join into the listings dataframe
listings_df <- calendar_df %>% 
          mutate(daily_revenue=daily_revenue(available, price.local_price)) %>% 
          group_by(id) %>% 
          summarize(currency=first(price.local_currency),
                    occupancy=occupancy(available),
                    revenue_per_available_room=revenue_per_available_room(
                      available, sum(daily_revenue)),
                    average_offered_rate=mean(price.local_price),
                    average_daily_rate=average_daily_rate(available, 
                                                          sum(daily_revenue)),
                    yearly_revenue=yearly_revenue(available, sum(daily_revenue))
                    ) %>%
          right_join(listings_df)  
#I am removing some unused columns
listings_df <-  listings_df[ , -which(names(listings_df) %in% 
                                        c("picture_url", "currency"))]


listings_df$picture_count_bucket <- cut(listings_df$picture_count, breaks = 
                                          c(0,5,10,15,20,40,60,80,100000)) 


listings_df$reviews_count_bucket <- cut(listings_df$reviews_count, breaks = 
                                          c(0,5,10,15,20,40,60,80,100000)) 


listings_df$rate_category = cut(listings_df$average_daily_rate, 
                                        breaks =c(0,30,70,120,200,100000), 
                                        labels = 
                                c('cheap','budget','midclass','upper','luxury')) 

```


# Univariate Plots Section

## Yearly revenue (Outcome variable)

As I am interesting in the factors which influence the Yearly Revenue, 'yearly_revenue' is the outcome variable.

```{r echo=FALSE, warning=FALSE, message=FALSE, fig.width=12, fig.height=6, Univariate_Plots}

p1 <- qplot(listings_df$yearly_revenue)
p2 <- qplot(factor(''), listings_df$yearly_revenue, geom='boxplot')
  
p3 <- ggplot(data=listings_df, mapping=aes(x=yearly_revenue)) +
  geom_histogram(binwidth = 0.1, alpha=0.6, color='white') + 
  geom_density(aes(y=..count../10), color='red') +
  scale_x_log10(breaks=c(100, 200, 500, 1000, 2000, 5000, 10000, 20000, 50000, 100000), 
                labels=c('0.1k','0.2k','0.5k','1k','2k','5k','10k','20k','50k','100k')
                ) +
  coord_cartesian(xlim=c(100, 100000))


p4 <- ggplot(data=listings_df, mapping=aes(y=yearly_revenue,x=factor(''))) + 
  geom_boxplot() + 
  coord_cartesian(ylim=c(0,15000))
  
grid.arrange(p1, p2, p3, p4, nrow = 2)

```

In the first row I plotted the histogram and a boxplot of yearly revenue unpolished. 

It is a positively skewed long tailed graph, you can also see some extreme outliers with
a expected yearly income of over EUR100,000.

In the second row I did a log tranformation to get a nearly gaussian curve. 

The second boxplot visualizes the median, lower and upper quartile of the yearly revenue by changing 
the limits of the plot.

## Average Daily Rate

```{r echo=FALSE, warning=FALSE, message=FALSE, fig.width=12, fig.height=10, Univariate_Plots_ADR}
adr <- qplot(subset(listings_df, average_daily_rate > 0)$average_daily_rate)

adr_mag <- qplot(listings_df$average_daily_rate, binwidth=3) +
              geom_density(aes(y=..count../0.25), color='red') +
              coord_cartesian(xlim=c(15,250),ylim=c(0,75))
adr_box <- qplot(c(''), listings_df$average_daily_rate, geom='boxplot')
adr_box_magn <- qplot(c(''), listings_df$average_daily_rate, geom='boxplot') +
            coord_cartesian(ylim=c(10,90))

rate_category <- qplot(listings_df$rate_category)

grid.arrange(adr, adr_box, adr_mag, adr_box_magn, rate_category, nrow=3)

```

The graph of the daily price is a long tailed. I zoomed in into the
rates between 0 and 250 to better see what's going on in this range.

I categorized the rates into 5 categories to get another view onto the data.

The most listings are in the lower rate segment. You can see this as well from the histogram for the rate category.

The box plot are illustrating the median, upper and lower quartile. The median average daily rate is about EUR 37 the lower quartile is at about 20 the upper quartile at about 80.




## Number of beds, bedrooms, bathrooms, person capacity

```{r echo=FALSE, warning=FALSE, message=FALSE, fig.width=12, fig.height=6, Univariate_Plots_Bedrooms}

beds <- qplot(as.factor(listings_df$beds))
bedrooms <- qplot(as.factor(listings_df$bedrooms))
bathrooms <- qplot(as.factor(listings_df$bathrooms))
person_capacity <- qplot(as.factor(listings_df$person_capacity))

grid.arrange(beds, bedrooms, bathrooms,person_capacity, nrow=2)
```

I visualized the variables 'beds', 'bedrooms', 'bathrooms' and person_capacity as bar charts by transforming
these continuous variables into categorical variables.

The values of these variables are very limited so they can be treated as categorical.

These variables are related to each other as all persons need a bed and beds are located in bedrooms. If unit has a higher person capacity it has most probably more beds, bedrooms  and bathrooms.

What you can see from data is that the majority of units has a person capacity of 2 and has 1 bedroom with 1 bed and 1 bathroom.



## Instant bookability, Preferation of families, Star Rating, Room type and Rate category 


```{r echo=FALSE, warning=FALSE, message=FALSE, fig.width=12, fig.height=6}
instant_bookable <- qplot(listings_df$instant_bookable)
is_family_preferred <- qplot(listings_df$is_family_preferred)
star_rating <- qplot(as.factor(listings_df$star_rating))
room_type <- qplot(listings_df$room_type)
grid.arrange(instant_bookable, is_family_preferred, 
             star_rating, room_type, nrow=2)
```

Most of hosts prefer the option that their listings are instant bookable, as they don't want to deal with too many inquiries and the guests likethat they are able to finalize the booking immediatly.

Most of the listings are not prefering families. These might be due the fact, that they simple don't have the capacity to host a family. Remember that most of the listings just have a person capacity of 2.

The majority of the listings have rating of 4.5 or 5 star or no rating at all. The reason for the good ratings is that Airbnb normally kicks you, if your rating is lower than 4. There are many listings without a rating yet. These listings are never booked or quiet new.

Most of the listings are either an entire home or a private room. There hardly any shared rooms on Bali.





## Picture count


```{r echo=FALSE, warning=FALSE, message=FALSE,fig.width=12, fig.height=8}

p1 <- qplot(listings_df$picture_count)
p2 <- qplot(as.factor('0'),listings_df$picture_count, geom='boxplot')
p3 <- ggplot(data=listings_df, mapping=aes(x=picture_count)) +
      geom_histogram(binwidth = 0.48, alpha=0.6, color='white')+
      geom_density(aes(y=..count../2), color='red') +
      scale_x_sqrt(breaks=c(0,10,20,30,40,50)) +
      coord_cartesian(xlim=c(5,50))
p4 <- ggplot(data=listings_df, mapping=aes(x=as.factor('0'), y=picture_count)) +
      geom_boxplot() +
      coord_cartesian(ylim=c(0,40))
p6 <- qplot(listings_df$picture_count_bucket)


grid.arrange(p1, p2, p6, p3,
             p4, nrow=2)
```

In the first row I plotted the picture count as histogram and a boxplot unpolished.

As the picture count is a bit positively skewed I applied a square root transformation to get a gaussian distribution the graph. The second boxplot illustrates the median, the lower and upper quartile.

The most listings have between 10-30 pictures with a median of about 20. There is a peak in the picture count bucket bar chart for a number between 20 and 40.

This is due the reason that Airbnb reccomends to have about 20-30 photos. This is also the number of photos what a professional photographer from Airbnb  (you are getting it for free) is shooting for you.



## Reviews count

```{r echo=FALSE, warning=FALSE, message=FALSE, fig.width=12, fig.height=8}
p1 <- qplot(listings_df$reviews_count) + 
        scale_x_continuous(breaks=seq(0,150,20))
p2 <- qplot(as.factor('0'),listings_df$reviews_count, geom='boxplot')
p3 <- qplot(listings_df$reviews_count, binwidth=.64) + 
        scale_x_log10(breaks=seq(0,100,10)) +
        coord_cartesian(xlim=c(5,70))
p4 <- ggplot(data=listings_df, mapping=aes(x=as.factor('0'), y=reviews_count)) +
        geom_boxplot() +
        coord_cartesian(ylim=c(0,40))
p5 <- qplot(listings_df$reviews_count_bucket)


grid.arrange(p1, p2, p5, p3, p4, nrow=2)
```




First I plotted the reviews count unpolished.

The distribution is very long tailed. Therefore I applied a log10 transformation with binwidth of 0.65 we are able to transform into a kind of normal distribution.

I visualized the distribution as well as boxplots. The median of reviews count is 10. The 25% quartile at about 2 and the 75% quartile at 31. There are outliers with reviews count over 75.

On the bar chart with the review count buckets we can see three peaks. The first is at the bucket between 0-5 reviews, the second one at the bucket with 20-30 and the last one at the bucket with no data at all.



# Univariate Analysis



## Structure of Data Set


There are two datsets used within this analysis: One dataset with the meta data of
1236 Airbnb listings and a second one with the booking calendar for 189 days of each listing. 

I aggregated some additional variables from the second dataset: 

   - Occupancy: Number of days a listing was not available divided by total number of days
   - Average Daily Rate: The average of all rates guests were paying for a place
   - Revenue per available room: Average Daily rate times occupancy
   - Yearly Revenue: Revenue per available room times 365. Estimation of revenue in a year.
   
I also created three additional categorical variables for the picture and review count by creating buckets to better see trends for these continous variables. The third variable I created categorizes the listings into different rate categories.
   


| Column        | Description           | Scale | Category values |
| ------------- |---------------------- | ----- | --------------- |
| id | Airbnb listing ID | unique key |  |
| name | Title of Airbnb Listing | unique key |  |
| yearly_revenue | Estimated Yearly revenue of a listing based on revenue in period |continuous |  |
| occupancy | Occupancy of listing in period |continuous |  |
| average_daily_rate | Average rate per day a guest paid in period |continuous |  |
| rate_category | Rate category based on average daily rate | ordinal  |'cheap','budget','midclass','upper','luxury'  |
| bathrooms | Number of bathrooms | continuous |  |
| bedrooms | Number of bedrooms | continuous |  |
| beds | Number of beds | continuous |  |
| person_capacity| Person capacity | continuous|  |
| instant_bookable| Whether a guest has to inquire to book or not | nominal | 'true', 'false |
| is_family_preferred| Whether families are preferred or not | nominal | 'true', 'false |
| picture_count | Number of pictures | continuous |  |
| picture_count_bucket | Buckets for different numbers of pictures | nominal | '-10', '10-20'... '> 80'|
| reviews_count | Number of reviews | continuous |  |
| reviews_count_bucket | Buckets for different numbers of reviews | nominal | '-10', '10-20'... '> 80'|
| room_type | Type of room | nominal | 'entire home', 'private room', 'shared room'|
| star_rating | Rating of the room | ordinal | '1.0', '1.5', '2.0', '2.5', '3.0', '3.5', '4.0', '4.5', '5.0' |


```{r}
str(listings_df)
```

I merged the aggregated data into the original dataset to get the dataset I am using throughout this analysis. It includes 1236 oberservations on 19 variables.

As a Airbnb host you are mainly interested in the revenue you can earn with a place. Therefore this variable is the outcome variable. Goal is to identify the features that influence this outcome variable and develop a prediction model.

There variables in the dataset that definitely influence the yearly revenue. There must be relationship between the number of bedrooms, bathrooms, beds and person capacity. It is interesting to analyse this deeper.

I am interested how big the influence of the different variables are.

It would be also interesting to group some of the categories together, as they are strongly related
to each other. As said before number of bedrooms, bathrooms and beds are stronglys related to person_capacity as well the room type.

I did a log transformation on the yearly revenue as it is a positively skewed long tail graph to transform it into a more gaussian curve to better analyse the curve. I also did transformations 
on the picture_count (Square root) and on reviews_count (log10)  for the same reasons.


# Bivariate Plots Section

I am plotting  a correlation of all variables to get a first impression of the relations between the variables:

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.width=12, fig.height=10}
ggcorr(listings_df[ , -which(names(listings_df) %in% c("id"))], 
       palette = "RdYlGn", label=TRUE)
```

First I am interested in the features that influence the yearly revenue. There is a correlation coefficent of 1 between the yearly revenue and the revenue per available room, but you have to keep in mind that the yearly revenue is calculated with the revenue per available room. So this correlation is "natural". There is a strong correlation between the average daily rate and yearly revenue as well with the occupancy and the yearly revenue. This is what you expect: The more you charge the higher the revenue and the more days a place is occupied the more you earn. But a higher rate might also drop the occupancy, especially if your place is overpriced in compare to similiar places. 

There is correlation between the average daily rate and the occupancy, but it is not as strong as you might expect. 

The number of pictures and number of reviews are having a weak relationship to the yearly revenue. But it has positive influence onto the yearly revenue to have more pciture and reviews.

You can see a strong relationship between number of bedrooms, bathrooms, beds and person_capacity as they all kind of metrics which describe the size of a place. This is also what you expect: Bigger places can charge more. 

The Occupancy has a weak relationship with the reviews_count and picture_count: More Pictures and more reviews is resulting in a slightly higher occupancy.


## Person Capacity, Number of bathrooms, bedrooms and beds




```{r echo=FALSE, message=FALSE, warning=FALSE, fig.width=12, fig.height=6, Bivariate_Plots}
listings_df[ , +which(names(listings_df) %in% 
                        c('yearly_revenue','person_capacity', 'beds', 
                                        'bedrooms','bathrooms',
                                        'star_rating'))] %>%
  gather(-yearly_revenue, key = "var", value = "value")  %>%
  mutate(value = as.numeric(value)) %>%
  ggplot(aes(y = yearly_revenue, x = value)) +
    geom_point(alpha=1/12, position = 'jitter') +
    scale_y_log10(breaks=c(500, 750, 1000, 2500, 5000, 10000, 20000, 30000, 50000)) +
    facet_wrap(~ var, scales = "free") 
```


```{r echo=FALSE, message=FALSE, warning=FALSE, fig.width=12, fig.height=6, Bivariate_Plots}
listings_df[ , +which(names(listings_df) %in% 
                        c('yearly_revenue','person_capacity', 'beds', 
                                        'bedrooms','bathrooms',
                                        'star_rating'))] %>%
  gather(-yearly_revenue, key = "var", value = "value")  %>%
  mutate(value = as.factor(as.numeric(value))) %>%
  ggplot(aes(y = yearly_revenue, x = value)) +
    geom_boxplot() +
    scale_y_log10(breaks=c(500, 750, 1000, 2500, 5000, 10000, 20000, 30000, 50000)) +
    facet_wrap(~ var, scales = "free") 
```

You can see that the more bathrooms a place has the more you can earn up to a certain limit. This can be seen in the scatter plot and on the boxplot. There not many places with more than 3 bathrooms. Therefore you cannot say anything for these places.

You can see that the yearly revenue is higher the more bedrooms a listing has up to a limit of 4. Over 4 there is not enough data.

You can see that the yearly revenue is higher the more beds a listing has up to a limit of 4. Over 4 there is not enough data.

The higher the person capacity the higher is the yearly revenue. The limit here is 6-8. We cannot say too much about places over 8.

I have to keep these limits in mind for the validity of our prediction model. 

Something else what I noticed is that, it is better to have a place with a even number.

The better your star rating the better is your revenue. Having no rating at all is better than having a rating of 4 and lower.



```{r echo=FALSE, message=FALSE, warning=FALSE}
listings_df[ , -which(names(listings_df) %in% 
                        c("average_daily_rate", "id", "name", 
                          "occupancy", "revenue_per_available_room", 
                          "picture_count", "instant_bookable", "bedrooms", 
                          "bathrooms", "beds", "person_capacity", 
                          "reviews_count", "star_rating"))] %>%
  gather(-yearly_revenue, key = "var", value = "value") %>%
  ggplot(aes(y = yearly_revenue, x = value)) +
    geom_boxplot() +
    scale_y_log10(breaks=c(500, 750, 1000, 2500, 5000, 7500, 10000, 20000, 30000, 50000)) +
    facet_wrap(~ var, scales = "free") 
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
listings_df[ , +which(names(listings_df) %in% 
                        c("yearly_revenue","instant_bookable", "room_type"))] %>%
  gather(-yearly_revenue, key = "var", value = "value") %>%
  ggplot(aes(y = yearly_revenue, x = value)) +
    geom_boxplot() +
    scale_y_log10(breaks=c(500, 750, 1000, 2500, 5000, 7500, 10000, 20000, 30000, 50000)) +
    facet_wrap(~ var, scales = "free") 
```

```{r echo=FALSE, message=FALSE, warning=FALSE}


p1 <- ggplot(data=listings_df, mapping=aes(x=picture_count, y=yearly_revenue)) +
      geom_point(alpha=0.2, position = 'jitter') +
      scale_y_log10(breaks=c(500, 750, 1000, 2500, 5000, 7500, 10000, 20000, 30000, 50000)) +
      scale_x_log10() +
      geom_smooth(method='lm')   
p2 <- ggplot(data=listings_df, mapping=aes(x=picture_count_bucket, y=yearly_revenue)) +
      geom_boxplot() +
      scale_y_log10(breaks=c(500, 750, 1000, 2500, 5000, 7500, 10000, 20000, 30000, 50000)) +
      geom_smooth(method='lm')   


grid.arrange(p1, p2, nrow=1)
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
p1 <- ggplot(data=listings_df, mapping=aes(x=reviews_count, y=yearly_revenue)) +
      geom_point(alpha=0.2, position = 'jitter') +
      scale_y_log10(breaks=c(500, 750, 1000, 2500, 5000, 7500, 10000, 20000, 30000, 50000)) +
      scale_x_log10() +
      geom_smooth(method='lm')   
p2 <- ggplot(data=listings_df, mapping=aes(x=reviews_count_bucket, y=yearly_revenue)) +
      geom_boxplot() +
      scale_y_log10(breaks=c(500, 750, 1000, 2500, 5000, 7500, 10000, 20000, 30000, 50000)) +
      geom_smooth(method='lm')   


grid.arrange(p1, p2, nrow=1)
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
listing_prop_df <- listings_df[,+which(names(listings_df) %in% 
                                      c('yearly_revenue','person_capacity', 'beds', 
                                        'bedrooms','bathrooms',
                                        'picture_count', 'reviews_count',
                                        'star_rating',"instant_bookable", "room_type"
                                        ))]
ggcorr(listing_prop_df, 
       palette = "RdYlGn", label=TRUE)
```



# Bivariate Analysis

> **Tip**: As before, summarize what you found in your bivariate explorations
here. Use the questions below to guide your discussion.

### Talk about some of the relationships you observed in this part of the \
investigation. How did the feature(s) of interest vary with other features in \
the dataset?

### Did you observe any interesting relationships between the other features \
(not the main feature(s) of interest)?

### What was the strongest relationship you found?


# Multivariate Plots Section

> **Tip**: Now it's time to put everything together. Based on what you found in
the bivariate plots section, create a few multivariate plots to investigate
more complex interactions between variables. Make sure that the plots that you
create here are justified by the plots you explored in the previous section. If
you plan on creating any mathematical models, this is the section where you
will do that.

```{r echo=FALSE, Multivariate_Plots}

```

# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. Were there features that strengthened each other in terms of \
looking at your feature(s) of interest?

### Were there any interesting or surprising interactions between features?

### OPTIONAL: Did you create any models with your dataset? Discuss the \
strengths and limitations of your model.

------

# Final Plots and Summary

> **Tip**: You've done a lot of exploration and have built up an understanding
of the structure of and relationships between the variables in your dataset.
Here, you will select three plots from all of your previous exploration to
present here as a summary of some of your most interesting findings. Make sure
that you have refined your selected plots for good titling, axis labels (with
units), and good aesthetic choices (e.g. color, transparency). After each plot,
make sure you justify why you chose each plot by describing what it shows.

### Plot One
```{r echo=FALSE, Plot_One}

```

### Description One


### Plot Two
```{r echo=FALSE, Plot_Two}

```

### Description Two


### Plot Three
```{r echo=FALSE, Plot_Three}

```

### Description Three

------

# Reflection

> **Tip**: Here's the final step! Reflect on the exploration you performed and
the insights you found. What were some of the struggles that you went through?
What went well? What was surprising? Make sure you include an insight into
future work that could be done with the dataset.

> **Tip**: Don't forget to remove this, and the other **Tip** sections before
saving your final work and knitting the final report!
