---
title: Exploration of factors which influence the yearly revenue of an Airbnb listing
  on Bali by Jens Laufer
output:
  pdf_document: default
  html_notebook: default
  html_document: default
---

Everybody heard about Airbnb, which is the fastest growing holiday website in the world. 

I am Airbnb host since 2015 and indeed it was one of the factors that lead me into the fascinating world of Data Science as I want to improve my own listings and find new opportunities to invest.

Currently I am helping an agency on Bali, Indonesia to improve their listings on Bali, Indonesia to gain experience with data analysis. 

I want to answer the question what factors influence the yearly revenue of an Airbnb listing on Bali.

The dataset for this project is scraped from Airbnb through there internal REST Api with python scripts, which are not part of this analysis. It includes 1236 listings on Bali and the booking calendar ofthe calendar for the next 6 months with about 230,000 entries.




```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
library(ggplot2)
library(tidyr)
library(dplyr)
library(GGally)
library(gridExtra)
```

```{r echo=FALSE, message=FALSE, Load_the_Data}
listings_df <- read.csv('data/airbnb.listings.bali.csv')
calendar_df <- read.csv('data/airbnb.listing.calendar.dates.bali.csv')

```


```{r echo=FALSE, message=FALSE}
# I am transforming the available into a Boolean value
calendar_df$available = as.logical(calendar_df$available == "true")

```





```{r echo=FALSE, warning=FALSE, message=FALSE, Creation_of_high_level_dataframe}
# Based on the "low level" calendar data I am calculating several metrics, 
# which are important for the hotel industry  (http://www.traveltripper.com/blog/the-most-important-metrics-for-hotel-revenue-managers/)
# These metrics are:
# Occupancy (OCC): Number of days a listing is not available / Total number of days
# Revenue: The total revenue in the period
# Average daily rate (ADR): Total revenue / Number of days a listing is not available
# Revenue per availabe room (RevPAR): Total revenue / Total number of days
# Yearly Revenue (estimated): 365 x Revenue per available room

 daily_revenue <- function(available, price) {
    ifelse(available == FALSE, price, 0)
 }

 average_daily_rate <- function(available, revenue) {
   if(length(which(available==FALSE)) > 0){
      revenue / length(which(available==FALSE))
   }
   else{
      0
   }
 }

 occupancy <- function(available) {
    length(which(available==FALSE)) / length(available)
 }
 
 revenue_per_available_room <- function(available, revenue) {
    average_daily_rate(available, revenue) * occupancy(available)
 }
 
 yearly_revenue <- function(available, revenue) {
    365 * revenue_per_available_room(available,revenue)
 }
 
# Calculation of metrics and join into the listings dataframe
listings_df <- calendar_df %>% 
          mutate(daily_revenue=daily_revenue(available, price.local_price)) %>% 
          group_by(id) %>% 
          summarize(currency=first(price.local_currency),
                    occupancy=occupancy(available),
                    revenue_per_available_room=revenue_per_available_room(available, sum(daily_revenue)),
                    average_daily_rate=average_daily_rate(available, sum(daily_revenue)),
                    yearly_revenue=yearly_revenue(available, sum(daily_revenue))
                    ) %>%
          right_join(listings_df)  
listings_df$star_rating = as.factor(listings_df$star_rating)
#I am removing some unused columns
listings_df <-  listings_df[ , -which(names(listings_df) %in% c("picture_url", "currency"))]
```




## Structure of Data Set


| Column        | Description           | Scale | Category values |
| ------------- |---------------------- | ----- | --------------- |
| id | Airbnb listing ID | unique key |  |
| name | Title of Airbnb Listing | unique key |  |
| yearly_revenue | Estimated Yearly revenue of a listing based on revenue in period |continuous |  |
| occupancy | Occupancy of listing in period |continuous |  |
| average_daily_rate | Average rate per day a guest paid in period |continuous |  |
| bathrooms | Number of bathrooms | continuous |  |
| bedrooms | Number of bedrooms | continuous |  |
| beds | Number of beds | continuous |  |
| person_capacity| Person capacity | continuous|  |
| instant_bookable| Whether a guest has to inquire to book or not | nominal | 'true', 'false |
| is_family_preferred| Whether families are preferred or not | nominal | 'true', 'false |
| picture_count | Number of pictures | continuous |  |
| reviews_count | Number of reviews | continuous |  |
| room_type | Type of room | nominal | 'entire home', 'private room', 'shared room'|
| star_rating | Rating of the room | 'ordinal' | '1.0', '1.5', '2.0', '2.5', '3.0', '3.5', '4.0', '4.5', '5.0' |


```{r}
str(listings_df)
```

The dataset contains 1236 rows with 16 variables, 12 of them are continuous and 6 categorical. 



# Univariate Plots Section

## Yearly revenue (Outcome variable)

As I am interesting in the factors which influence the Yearly Revenue 'yearly_income' is the outcome variable.

```{r echo=FALSE, warning=FALSE, message=FALSE, Univariate_Plots}

p1 <- qplot(listings_df$yearly_revenue)
p2 <- qplot(factor(''), listings_df$yearly_revenue, geom='boxplot')
  
p3 <- ggplot(data=listings_df, mapping=aes(x=yearly_revenue)) +
  geom_histogram() + 
  scale_x_log10(breaks=c(100, 200, 500, 1000, 2000, 5000, 10000, 20000, 50000, 100000), 
                labels=c('0.1k','0.2k','0.5k','1k','2k','5k','10k','20k','50k','100k')
                ) +
  coord_cartesian(xlim=c(100, 100000))


p4 <- ggplot(data=listings_df, mapping=aes(y=yearly_revenue,x=factor(''))) + 
  geom_boxplot() + 
  coord_cartesian(ylim=c(0,15000))
  
grid.arrange(p1, p2, p3, p4, nrow = 2)

```

In the first row I plotted the histogram and a boxplot of yearly revenue unpolished. 

It is a positively skewed long tailed graph, we can also see some extreme outliers with
a expected yearly income of over ???100,000.. 

In the second row I did a log tranformation to get a nearly gaussian curve. 

The second boxplot visualizes the median, lower and upper quartile of the yearly revenue.



## Number of beds, bedrooms, bathrooms, person capacity

```{r echo=FALSE, warning=FALSE, message=FALSE, Univariate_Plots_Bedrooms}

beds <- qplot(as.factor(listings_df$beds))
bedrooms <- qplot(as.factor(listings_df$bedrooms))
bathrooms <- qplot(as.factor(listings_df$bathrooms))
person_capacity <- qplot(as.factor(listings_df$person_capacity))

grid.arrange(beds, bedrooms, bathrooms,person_capacity, nrow=2)
```

I visualized the variables 'beds', 'bedrooms', 'bathrooms' and person_capacity as bar charts by transforming
these continuous variables into categorical variables.

The values of these variables are very limited so they can treated as categorical.

These variables are somehow related to each other as all persons need a bed and beds are located in bedrooms. If unit has a higher person capacity it has most probably more beds, bedrooms  and bathrooms.

What we can see from data is that the majority of units has a person capacity of 2 and has 1 bedroom with 1 bed and 1 bathroom.


## Instant bookable, Preferation of families, Rating and Room type


```{r echo=FALSE, warning=FALSE, message=FALSE}
instant_bookable <- qplot(listings_df$instant_bookable)
is_family_preferred <- qplot(listings_df$is_family_preferred)
star_rating <- qplot(listings_df$star_rating)
room_type <- qplot(listings_df$room_type)
grid.arrange(instant_bookable, is_family_preferred, star_rating, room_type, nrow=2)
```

Most of the listings prefer the instant_bookable option, as the hosts don't want to deal with too many inquiries and the guests like that they are able to finalize the booking immediatly.

Most of the listings are not prefering families. These might be due the fact, that they simple don't have the capacity to host a family. Remember that most of the listings just have a person capacity of 2.

The majority of the listings have rating of 4.5 or 5 star or no rating at all. The reason for the good ratings is that Airbnb normally kicks you, if your rating is lower than 4. There are many listings without a rating yet. These listings are never booked or quiet new.

Most of the listings are either an entire home or a private room. There hardly any shared rooms on Bali.




## Picture count


```{r echo=FALSE, warning=FALSE, message=FALSE}

p1 <- qplot(listings_df$picture_count)
p2 <- qplot(as.factor('0'),listings_df$picture_count, geom='boxplot')
p3 <- ggplot(data=listings_df, mapping=aes(x=picture_count)) +
      geom_histogram(binwidth = 0.45)+
      scale_x_sqrt(breaks=c(0,10,20,30,40,50)) +
      coord_cartesian(xlim=c(5,50))
p4 <- ggplot(data=listings_df, mapping=aes(x=as.factor('0'), y=picture_count)) +
      geom_boxplot() +
      coord_cartesian(ylim=c(0,40))


grid.arrange(p1,p2,p3,p4,nrow=2)
```

In the first row I plotted the picture count as histogram and a boxplot unpolished.

As the picture count is a bit positively skewed I app,ied a square root transformation to normalize the graph. The second boxplot is to illustare the median, the lower and upper quartile.

The most listings have between 10-30 pictures with a median of about 20.



## Reviews count

```{r echo=FALSE, warning=FALSE, message=FALSE}
p1 <- qplot(listings_df$reviews_count) + 
        scale_x_continuous(breaks=seq(0,150,20))
p2 <- qplot(as.factor('0'),listings_df$reviews_count, geom='boxplot')
p3 <- qplot(listings_df$reviews_count, binwidth=1) + 
        scale_x_sqrt(breaks=seq(0,100,10)) +
        coord_cartesian(xlim=c(5,70))
p4 <- ggplot(data=listings_df, mapping=aes(x=as.factor('0'), y=reviews_count)) +
        geom_boxplot() +
        coord_cartesian(ylim=c(0,40))


grid.arrange(p1, p2, p3, p4, nrow=2)
```




First I plotted the reviews count unpolished.

The distribution is long tailed.







> **Tip**: Make sure that you leave a blank line between the start / end of
each code block and the end / start of your Markdown text so that it is
formatted nicely in the knitted text. Note as well that text on consecutive
lines is treated as a single space. Make sure you have a blank line between
your paragraphs so that they too are formatted for easy readability.

# Univariate Analysis

> **Tip**: Now that you've completed your univariate explorations, it's time to
reflect on and summarize what you've found. Use the questions below to help you
gather your observations and add your own if you have other thoughts!

### What is the structure of your dataset?

### What is/are the main feature(s) of interest in your dataset?

### What other features in the dataset do you think will help support your \
investigation into your feature(s) of interest?

### Did you create any new variables from existing variables in the dataset?

### Of the features you investigated, were there any unusual distributions? \
Did you perform any operations on the data to tidy, adjust, or change the form \
of the data? If so, why did you do this?


# Bivariate Plots Section

> **Tip**: Based on what you saw in the univariate plots, what relationships
between variables might be interesting to look at in this section? Don't limit
yourself to relationships between a main output feature and one of the
supporting variables. Try to look at relationships between supporting variables
as well.

```{r echo=FALSE, message=FALSE, warning=FALSE, Bivariate_Plots}
listings_df[ , +which(names(listings_df) %in% 
                        c('yearly_revenue','person_capacity', 'beds', 
                                        'bedrooms','bathrooms',
                                        'star_rating'))] %>%
  gather(-yearly_revenue, key = "var", value = "value")  %>%
  mutate(value = as.factor(as.numeric(value))) %>%
  ggplot(aes(y = yearly_revenue, x = value)) +
    geom_boxplot() +
    scale_y_log10(breaks=c(500, 750, 1000, 2500, 5000, 10000, 20000, 30000, 50000)) +
    facet_wrap(~ var, scales = "free") 
```
```{r echo=FALSE, message=FALSE, warning=FALSE}
listings_df[ , -which(names(listings_df) %in% 
                        c("average_daily_rate", "id", "name", 
                          "occupancy", "revenue_per_available_room", 
                          "picture_count", "instant_bookable", "bedrooms", 
                          "bathrooms", "beds", "person_capacity", 
                          "reviews_count", "star_rating"))] %>%
  gather(-yearly_revenue, key = "var", value = "value") %>%
  ggplot(aes(y = yearly_revenue, x = value)) +
    geom_boxplot() +
    scale_y_log10(breaks=c(500, 750, 1000, 2500, 5000, 7500, 10000, 20000, 30000, 50000)) +
    facet_wrap(~ var, scales = "free") 
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
listings_df[ , +which(names(listings_df) %in% 
                        c("yearly_revenue","instant_bookable", "room_type"))] %>%
  gather(-yearly_revenue, key = "var", value = "value") %>%
  ggplot(aes(y = yearly_revenue, x = value)) +
    geom_boxplot() +
    scale_y_log10(breaks=c(500, 750, 1000, 2500, 5000, 7500, 10000, 20000, 30000, 50000)) +
    facet_wrap(~ var, scales = "free") 
```

```{r echo=FALSE, message=FALSE, warning=FALSE}

listings_df$picture_count_bucket <- cut(listings_df$picture_count, breaks = 
                                          c(0,5,10,15,20,40,60,80,100)) 

p1 <- ggplot(data=listings_df, mapping=aes(x=picture_count, y=yearly_revenue)) +
      geom_point(alpha=0.2, position = 'jitter') +
      scale_y_log10(breaks=c(500, 750, 1000, 2500, 5000, 7500, 10000, 20000, 30000, 50000)) +
      geom_smooth(method='lm')   
p2 <- ggplot(data=listings_df, mapping=aes(x=picture_count_bucket, y=yearly_revenue)) +
      geom_boxplot() +
      scale_y_log10(breaks=c(500, 750, 1000, 2500, 5000, 7500, 10000, 20000, 30000, 50000)) +
      geom_smooth(method='lm')   


grid.arrange(p1, p2, nrow=1)
```

```{r echo=FALSE, message=FALSE, warning=FALSE}

listings_df$reviews_count_bucket <- cut(listings_df$reviews_count, breaks = 
                                          c(0,5,10,15,20,40,60,80,100)) 

p1 <- ggplot(data=listings_df, mapping=aes(x=reviews_count, y=yearly_revenue)) +
      geom_point(alpha=0.2, position = 'jitter') +
      scale_y_log10(breaks=c(500, 750, 1000, 2500, 5000, 7500, 10000, 20000, 30000, 50000)) +
      geom_smooth(method='lm')   
p2 <- ggplot(data=listings_df, mapping=aes(x=reviews_count_bucket, y=yearly_revenue)) +
      geom_boxplot() +
      scale_y_log10(breaks=c(500, 750, 1000, 2500, 5000, 7500, 10000, 20000, 30000, 50000)) +
      geom_smooth(method='lm')   


grid.arrange(p1, p2, nrow=1)
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
listing_prop_df <- listings_df[,+which(names(listings_df) %in% 
                                      c('yearly_revenue','person_capacity', 'beds', 
                                        'bedrooms','bathrooms',
                                        'picture_count', 'reviews_count',
                                        'star_rating',"instant_bookable", "room_type"
                                        ))]
ggcorr(listing_prop_df, 
       palette = "RdYlGn", label=TRUE)
```



# Bivariate Analysis

> **Tip**: As before, summarize what you found in your bivariate explorations
here. Use the questions below to guide your discussion.

### Talk about some of the relationships you observed in this part of the \
investigation. How did the feature(s) of interest vary with other features in \
the dataset?

### Did you observe any interesting relationships between the other features \
(not the main feature(s) of interest)?

### What was the strongest relationship you found?


# Multivariate Plots Section

> **Tip**: Now it's time to put everything together. Based on what you found in
the bivariate plots section, create a few multivariate plots to investigate
more complex interactions between variables. Make sure that the plots that you
create here are justified by the plots you explored in the previous section. If
you plan on creating any mathematical models, this is the section where you
will do that.

```{r echo=FALSE, Multivariate_Plots}

```

# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. Were there features that strengthened each other in terms of \
looking at your feature(s) of interest?

### Were there any interesting or surprising interactions between features?

### OPTIONAL: Did you create any models with your dataset? Discuss the \
strengths and limitations of your model.

------

# Final Plots and Summary

> **Tip**: You've done a lot of exploration and have built up an understanding
of the structure of and relationships between the variables in your dataset.
Here, you will select three plots from all of your previous exploration to
present here as a summary of some of your most interesting findings. Make sure
that you have refined your selected plots for good titling, axis labels (with
units), and good aesthetic choices (e.g. color, transparency). After each plot,
make sure you justify why you chose each plot by describing what it shows.

### Plot One
```{r echo=FALSE, Plot_One}

```

### Description One


### Plot Two
```{r echo=FALSE, Plot_Two}

```

### Description Two


### Plot Three
```{r echo=FALSE, Plot_Three}

```

### Description Three

------

# Reflection

> **Tip**: Here's the final step! Reflect on the exploration you performed and
the insights you found. What were some of the struggles that you went through?
What went well? What was surprising? Make sure you include an insight into
future work that could be done with the dataset.

> **Tip**: Don't forget to remove this, and the other **Tip** sections before
saving your final work and knitting the final report!
