---
output:
  pdf_document: default
  html_document: default
---

# Exploration of factors which influence the yearly revenue of an Airbnb listing on Bali by Jens Laufer

Everybody heard about Airbnb, which is the fastest growing holiday website in the world. 

I am Airbnb host since 2015 and indeed it was one of the factors that lead me into the fascinating world of Data Science as I want to improve my own listings and find new opportunities to invest.

Currently I am helping an agency on Bali, Indonesia to improve their listings on Bali, Indonesia to gain experience with data analysis. 

I want to answer the question what factors influence the yearly revenue of an Airbnb listing on Bali.

The dataset for this project is scraped from Airbnb through there internal REST Api with python scripts, which are not part of this analysis. It includes 1234 listings on Bali and the booking calendar for the next 6 months with about 230,000 entries.




```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
library(ggplot2)
library(tidyr)
library(dplyr)
library(GGally)
library(gridExtra)
```

```{r echo=FALSE, message=FALSE, Load_the_Data}
listings_df <- read.csv('data/airbnb.listings.bali.csv')
calendar_df <- read.csv('data/airbnb.listing.calendar.dates.bali.csv')

```


```{r echo=FALSE, message=FALSE}
# I am transforming the available into a Boolean value
calendar_df$available = as.logical(calendar_df$available == "true")

```





```{r echo=FALSE, warning=FALSE, message=FALSE, Creation_of_high_level_dataframe}
# Based on the "low level" calendar data I am calculating several metrics, 
# which are important for the hotel industry  (http://www.traveltripper.com/blog/the-most-important-metrics-for-hotel-revenue-managers/)
# These metrics are:
# Occupancy (OCC): Number of days a listing is not available / Total number of days
# Revenue: The total revenue in the period
# Average daily rate (ADR): Total revenue / Number of days a listing is not available
# Revenue per availabe room (RevPAR): Total revenue / Total number of days
# Yearly Revenue (estimated): 365 x Revenue per available room

 daily_revenue <- function(available, price) {
    ifelse(available == FALSE, price, 0)
 }

 average_daily_rate <- function(available, revenue) {
   if(length(which(available==FALSE)) > 0){
      revenue / length(which(available==FALSE))
   }
   else{
      0
   }
 }

 occupancy <- function(available) {
    length(which(available==FALSE)) / length(available)
 }
 
 revenue_per_available_room <- function(available, revenue) {
    average_daily_rate(available, revenue) * occupancy(available)
 }
 
 yearly_revenue <- function(available, revenue) {
    365 * revenue_per_available_room(available,revenue)
 }
 
# Calculation of metrics and join into the listings dataframe
listings_df <- calendar_df %>% 
          mutate(daily_revenue=daily_revenue(available, price.local_price)) %>% 
          group_by(id) %>% 
          summarize(currency=first(price.local_currency),
                    occupancy=occupancy(available),
                    revenue_per_available_room=revenue_per_available_room(available, sum(daily_revenue)),
                    average_daily_rate=average_daily_rate(available, sum(daily_revenue)),
                    yearly_revenue=yearly_revenue(available, sum(daily_revenue))
                    ) %>%
          right_join(listings_df)  
listings_df$star_rating = as.factor(listings_df$star_rating)
listings_df <-  listings_df[ , -which(names(listings_df) %in% c("picture_url", "currency"))]
```




## Structure of Data Set


| Column        | Description           | Scale | Category values |
| ------------- |---------------------- | ----- | --------------- |
| yearly_revenue | Estimated Yearly renenue based on revenue in period |continuous |  |
| bathrooms | Number of bathrooms | continuous |  |
| beds | Number of beds | continuous |  |
| instant_bookable| Whether a guest has to inquire to book or not | nominal | 'true', 'false |
| is_family_preferred| Whether families are preferred or not | nominal | 'true', 'false |
| person_capacity| Person capacity | continuous|  |
| picture_count | Number of pictures | continuous |  |
| reviews_count | Number of reviews | continuous |  |
| room_type | Type of room | nominal | 'entire home', 'private room', 'shared room'|
| star_rating | Rating of the room | 'ordinal' | '1.0', '1.5', '2.0', '2.5', '3.0', '3.5', '4.0', '4.5', '5.0' |


```{r}
str(listings_df)
```

The dataset contains 1236 rows with 16 variables, 12 of them are continuous and 4 categorical. 







> **Tip**: Before you create any plots, it is a good idea to provide a short
introduction into the dataset that you are planning to explore. Replace this
quoted text with that general information!

# Univariate Plots Section

> **Tip**: In this section, you should perform some preliminary exploration of
your dataset. Run some summaries of the data and create univariate plots to
understand the structure of the individual variables in your dataset. Don't
forget to add a comment after each plot or closely-related group of plots!
There should be multiple code chunks and text sections; the first one below is
just to help you get started.

```{r echo=FALSE, warning=FALSE, message=FALSE, Univariate_Plots}

p1 <- qplot(listings_df$yearly_revenue, geom='histogram')
p2 <- qplot(factor(''), listings_df$yearly_revenue, geom='boxplot')
  
p3 <- ggplot(data=listings_df, mapping=aes(x=yearly_revenue)) +
  geom_histogram() + scale_x_log10(breaks=c(0,100,500,1000,2000,5000,7500,10000,15000,20000))
p4 <- ggplot(data=listings_df, mapping=aes(y=yearly_revenue,x=factor(''))) + 
  geom_boxplot() + 
  coord_cartesian(ylim = c(0,15000))
  
grid.arrange(p1, p2, p3, p4, nrow = 2)
```

We have a positively skewed graph for the yearly revenue, we can also see some outliers with
a expected yearly income of over ??? 100,000 on the one site and very low income (< ???2000) on the other site. This needs ti ```



```{r echo=FALSE, warning=FALSE, message=FALSE, Univariate_Plots_Bedrooms}

beds <- qplot(as.factor(listings_df$beds))
bedrooms <- qplot(as.factor(listings_df$bedrooms))
bathrooms <- qplot(as.factor(listings_df$bathrooms))
person_capacity <- qplot(as.factor(listings_df$person_capacity))

grid.arrange(beds, bedrooms, bathrooms,person_capacity, nrow=2)
```




```{r echo=FALSE, warning=FALSE, message=FALSE}
instant_bookable <- qplot(listings_df$instant_bookable)
is_family_preferred <- qplot(listings_df$is_family_preferred)
star_rating <- qplot(listings_df$star_rating)
room_type <- qplot(listings_df$room_type)
grid.arrange(instant_bookable, is_family_preferred, star_rating, room_type, nrow=2)
```
```{r echo=FALSE, warning=FALSE, message=FALSE}

p1 <- qplot(listings_df$picture_count)
p2 <- qplot(as.factor('0'),listings_df$picture_count, geom='boxplot')
p3 <- ggplot(data=listings_df, mapping=aes(x=picture_count)) +
      geom_histogram()+
      scale_x_sqrt()
p4 <- ggplot(data=listings_df, mapping=aes(x=as.factor('0'), y=picture_count)) +
      geom_boxplot() +
      coord_cartesian(ylim=c(0,50))


grid.arrange(p1,p2,p3,p4,nrow=2)
```

```{r echo=FALSE, warning=FALSE, message=FALSE}
p1 <- qplot(listings_df$reviews_count)
p2 <- qplot(as.factor('0'),listings_df$reviews_count, geom='boxplot')
p3 <- ggplot(data=listings_df, mapping=aes(x=reviews_count)) +
      geom_histogram(binwidth = 0.5)+
      scale_x_sqrt()
p4 <- ggplot(data=listings_df, mapping=aes(x=as.factor('0'), y=reviews_count)) +
      geom_boxplot() +
      coord_cartesian(ylim=c(0,40))


grid.arrange(p1,p2,p3,p4,nrow=2)
```




The majority of listing have 1 bedroom, but there are listings with 0 bedrooms, which seems to be unrealistic on the other site there a few listings with more than 4 bedrooms, which must be a massive house.

Most of the listings has 1-6 beds. There must be a problem in the data: We had listings with more than 10 bedrooms but the maximum number of beds is 12.



> **Tip**: Make sure that you leave a blank line between the start / end of
each code block and the end / start of your Markdown text so that it is
formatted nicely in the knitted text. Note as well that text on consecutive
lines is treated as a single space. Make sure you have a blank line between
your paragraphs so that they too are formatted for easy readability.

# Univariate Analysis

> **Tip**: Now that you've completed your univariate explorations, it's time to
reflect on and summarize what you've found. Use the questions below to help you
gather your observations and add your own if you have other thoughts!

### What is the structure of your dataset?

### What is/are the main feature(s) of interest in your dataset?

### What other features in the dataset do you think will help support your \
investigation into your feature(s) of interest?

### Did you create any new variables from existing variables in the dataset?

### Of the features you investigated, were there any unusual distributions? \
Did you perform any operations on the data to tidy, adjust, or change the form \
of the data? If so, why did you do this?


# Bivariate Plots Section

> **Tip**: Based on what you saw in the univariate plots, what relationships
between variables might be interesting to look at in this section? Don't limit
yourself to relationships between a main output feature and one of the
supporting variables. Try to look at relationships between supporting variables
as well.

```{r echo=FALSE, message=FALSE, warning=FALSE, Bivariate_Plots}
listings_df[ , +which(names(listings_df) %in% 
                        c('yearly_revenue','person_capacity', 'beds', 
                                        'bedrooms','bathrooms',
                                        'star_rating'))] %>%
  gather(-yearly_revenue, key = "var", value = "value")  %>%
  mutate(value = as.factor(as.numeric(value))) %>%
  ggplot(aes(y = yearly_revenue, x = value)) +
    geom_boxplot() +
    scale_y_log10(breaks=c(500, 750, 1000, 2500, 5000, 10000, 20000, 30000, 50000)) +
    facet_wrap(~ var, scales = "free") 
```
```{r echo=FALSE, message=FALSE, warning=FALSE}
listings_df[ , -which(names(listings_df) %in% 
                        c("average_daily_rate", "id", "name", 
                          "occupancy", "revenue_per_available_room", 
                          "picture_count", "instant_bookable", "bedrooms", 
                          "bathrooms", "beds", "person_capacity", 
                          "reviews_count", "star_rating"))] %>%
  gather(-yearly_revenue, key = "var", value = "value") %>%
  ggplot(aes(y = yearly_revenue, x = value)) +
    geom_boxplot() +
    scale_y_log10(breaks=c(500, 750, 1000, 2500, 5000, 7500, 10000, 20000, 30000, 50000)) +
    facet_wrap(~ var, scales = "free") 
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
listings_df[ , +which(names(listings_df) %in% 
                        c("yearly_revenue","instant_bookable", "room_type"))] %>%
  gather(-yearly_revenue, key = "var", value = "value") %>%
  ggplot(aes(y = yearly_revenue, x = value)) +
    geom_boxplot() +
    scale_y_log10(breaks=c(500, 750, 1000, 2500, 5000, 7500, 10000, 20000, 30000, 50000)) +
    facet_wrap(~ var, scales = "free") 
```

```{r echo=FALSE, message=FALSE, warning=FALSE}

listings_df$picture_count_bucket <- cut(listings_df$picture_count, breaks = 
                                          c(0,5,10,15,20,40,60,80,100)) 

p1 <- ggplot(data=listings_df, mapping=aes(x=picture_count, y=yearly_revenue)) +
      geom_point(alpha=0.2, position = 'jitter') +
      scale_y_log10(breaks=c(500, 750, 1000, 2500, 5000, 7500, 10000, 20000, 30000, 50000)) +
      geom_smooth(method='lm')   
p2 <- ggplot(data=listings_df, mapping=aes(x=picture_count_bucket, y=yearly_revenue)) +
      geom_boxplot() +
      scale_y_log10(breaks=c(500, 750, 1000, 2500, 5000, 7500, 10000, 20000, 30000, 50000)) +
      geom_smooth(method='lm')   


grid.arrange(p1, p2, nrow=1)
```

```{r echo=FALSE, message=FALSE, warning=FALSE}

listings_df$reviews_count_bucket <- cut(listings_df$reviews_count, breaks = 
                                          c(0,5,10,15,20,40,60,80,100)) 

p1 <- ggplot(data=listings_df, mapping=aes(x=reviews_count, y=yearly_revenue)) +
      geom_point(alpha=0.2, position = 'jitter') +
      scale_y_log10(breaks=c(500, 750, 1000, 2500, 5000, 7500, 10000, 20000, 30000, 50000)) +
      geom_smooth(method='lm')   
p2 <- ggplot(data=listings_df, mapping=aes(x=reviews_count_bucket, y=yearly_revenue)) +
      geom_boxplot() +
      scale_y_log10(breaks=c(500, 750, 1000, 2500, 5000, 7500, 10000, 20000, 30000, 50000)) +
      geom_smooth(method='lm')   


grid.arrange(p1, p2, nrow=1)
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
listing_prop_df <- listings_df[,+which(names(listings_df) %in% 
                                      c('yearly_revenue','person_capacity', 'beds', 
                                        'bedrooms','bathrooms',
                                        'picture_count', 'reviews_count',
                                        'star_rating',"instant_bookable", "room_type"
                                        ))]
ggcorr(listing_prop_df, 
       palette = "RdYlGn", label=TRUE)
```



# Bivariate Analysis

> **Tip**: As before, summarize what you found in your bivariate explorations
here. Use the questions below to guide your discussion.

### Talk about some of the relationships you observed in this part of the \
investigation. How did the feature(s) of interest vary with other features in \
the dataset?

### Did you observe any interesting relationships between the other features \
(not the main feature(s) of interest)?

### What was the strongest relationship you found?


# Multivariate Plots Section

> **Tip**: Now it's time to put everything together. Based on what you found in
the bivariate plots section, create a few multivariate plots to investigate
more complex interactions between variables. Make sure that the plots that you
create here are justified by the plots you explored in the previous section. If
you plan on creating any mathematical models, this is the section where you
will do that.

```{r echo=FALSE, Multivariate_Plots}

```

# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. Were there features that strengthened each other in terms of \
looking at your feature(s) of interest?

### Were there any interesting or surprising interactions between features?

### OPTIONAL: Did you create any models with your dataset? Discuss the \
strengths and limitations of your model.

------

# Final Plots and Summary

> **Tip**: You've done a lot of exploration and have built up an understanding
of the structure of and relationships between the variables in your dataset.
Here, you will select three plots from all of your previous exploration to
present here as a summary of some of your most interesting findings. Make sure
that you have refined your selected plots for good titling, axis labels (with
units), and good aesthetic choices (e.g. color, transparency). After each plot,
make sure you justify why you chose each plot by describing what it shows.

### Plot One
```{r echo=FALSE, Plot_One}

```

### Description One


### Plot Two
```{r echo=FALSE, Plot_Two}

```

### Description Two


### Plot Three
```{r echo=FALSE, Plot_Three}

```

### Description Three

------

# Reflection

> **Tip**: Here's the final step! Reflect on the exploration you performed and
the insights you found. What were some of the struggles that you went through?
What went well? What was surprising? Make sure you include an insight into
future work that could be done with the dataset.

> **Tip**: Don't forget to remove this, and the other **Tip** sections before
saving your final work and knitting the final report!
