---
title: "Identifying Fraud from Enron Emails and Financial Data"
output:
  pdf_document: default
  html_notebook: default
  html_document: default
---



```{r echo=FALSE, message=FALSE, warning=FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(gridExtra)
library(knitr)
library(kableExtra)
library(stringr)
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
raw_df = read.csv('raw_data.csv')
without_nans_df = read.csv('cleaned_data1.csv')
cleaned_df = read.csv('cleaned_data2.csv')
kbest_results_df = read.csv('kbest.csv')
metrics_df = read.csv('metrics.csv')

nulls <- raw_df %>%
  gather(-name, -email_address, -poi, key = "var", value = "val")  %>%
  group_by(var) %>%
  summarise(number = dim(raw_df)[1] - sum(val == 'NaN') -1) %>%
  arrange(desc(number))
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
outlier_plots <- function(df){
df %>%
  gather(-name, -email_address, -poi, key = "var", value = "value")  %>%
  group_by(var) %>%
  mutate(max = max(value)) %>%
  ggplot(aes(x =poi, y=value)) +
   # geom_point(position='jitter', alpha=0.2, color=c('blue')) +
    geom_boxplot() +
    geom_text(aes(label=ifelse(value > .6 * max, as.character(name), '')))+
    facet_wrap(~var, scales = "free")
  
}
```



by Jens Laufer, 02/08/2017



## Introduction


Enron was formed in 1985 by a merge of Houston Natural Gas Co and InterNorth 
Inc. Kenneth Lay became Enron's first CEO and Chairman and transformed the 
company into a energy trader and supplier. Deregulation of the energy market and 
an aggressive company policy fired the growth of the company. By 2000 the 
company was one of the biggest companies in the US and a Wallstreet darling. 
Behind the scenes financial losses of the company were hidden behind a network 
of companies with the help of the CEOs Kenneth Lay and Jeff Skilling and the 
CFO Andrew Fastow and others. 
By the end of 2001 the company declared bankruptcy all of a 
sudden to the public. 
The U.S. Securities and Exchange Commission (SEC) began an investigation and 
unhided the biggest case of account fraud in the US history.
Kenneth Lay, Jeff Skilling, Andrew Fastow and several other excutives were later 
sentenced to prison.

The goal of this study is to develop a machine learning model to classify, if a
person is a so called person of interest (POI), who was involved in fraud.

For the study a dataset with email and financial data of 146 excutives is used. 
The original dataset can be downloaded from 
(https://www.cs.cmu.edu/enron/enron_mail_20150507.tgz) and was enriched with  
financial data.


## Data Exploration

The data set consists of `r dim(cleaned_df)[1]` observations and 
`r dim(cleaned_df)[2]-1` features.

There are 
`r dim(subset(cleaned_df, poi == 'True'))[1]` POI
and 
`r dim(subset(cleaned_df, poi == 'False'))[1]` non POIs in the dataset.



### Cleaning the Data



```{r echo=FALSE, message=FALSE, warning=FALSE}
nulls %>%
  ggplot(aes(x=var, y=number)) +
  geom_bar(stat='identity', alpha=0.5, fill=c('blue')) +
  scale_x_discrete(limits=rev(as.factor(nulls$var))) +
  labs(y='Number', x='Feature', title='Number of available features') +
  coord_flip()
```




Occurences of NaN values were checked. It can be seen that for some 
features not that many data points are available. Especially 'loan_advances'
just have 3 data points. This has to be kept in mind for the feature 
selection.

For financial data the NaN values were replaced with 0.


### Outliers


Outliers need to be removed from the dataset. As fraud detection is 
the goal of the study this has to be done with cautious as some POIs have 
outliers in their financial data. These outliers are
essential and cannot be thrown away as they characterize a POI. 
So just undesired outliers can be removed.

```{r echo=FALSE, fig.height=8, fig.width=16, message=FALSE, warning=FALSE}
outlier_plots(without_nans_df) 
```

From the figure we can see a outlier for the POI 'TOTAL'. These seems more a sum rather than a person.

The following outliers were removed:

'TOTAL' is removed, as it is the sum of a financial
feature data points.

'THE TRAVEL AGENCY IN THE PARK' is removed as it is not a person.

'LOCKHART EUGENE E' is removed as all features are NaN.

'KAMINSKI WINCENTY J': Non POI with outlier on 'from_messages'

'BHATNAGAR SANJAY': Non POI with outlier on 'from_messages'

'FREVERT MARK A': Non POI with outlier on 'deferrall_payments'

'LAVORATO JOHN J': Non POI with outlier on 'bonus'

'MARTIN AMANDA K': Non POI with outlier on 'long_term_incentives'

'WHITE JR THOMAS E': Non POI with outlier on 'restricted_stock'

'KEAN STEVEN J': Non POI with outlier on 'from_messages'

'ECHOLS JOHN B': Non POI with outlier on 'long_term_incentives'




```{r echo=FALSE, fig.height=8, fig.width=16, message=FALSE, warning=FALSE}
outlier_plots(cleaned_df) 
```
The figure shows the distribution of the dataset after cleaning the 'bad' 
outliers. As mentioned before everything seem ok now, there are only "desired"
outliers.



## Feature Engineering


3 new features were created:

   - total_financial_benefits: Sum of 'salary', 'bonus', 'total_stock_value', 
   and 'exercised_stock_options'. 
   
   - message_to_poi_ratio: Ratio a person  sends
   messages to POI
   
   - message_from_poi_ratio:  Ratio a person receives  messages from POI

The message ratios were created as it seems possible that POIs communicate more likely with each other.

'total_financial_benefits' was created as it summarizes all financial benefits a person gets. 


Missing values of 'from_messages' and 'to_messages' were replaced with the mean of these fields over the whole data set to not tap into division by zero problems.

Feature scaling was performed after the creation of message fraction 
features and before the creation of 'total_financial_benefits' as some algorithms like logistic regression and SVM will not work properly without scaling.  It was done before the fraction calculation to avoid division by 0 problems.



## Feature Selection



```{r echo=FALSE, message=FALSE, warning=FALSE}
  kbest_arranged <- kbest_results_df %>%
                     arrange(desc(score))
  
  kbest_arranged %>%
  ggplot(aes(x=var, y=score)) +
  geom_bar(stat='identity', alpha=0.5, fill=c('red')) +
  scale_x_discrete(limits=rev(as.factor(kbest_arranged$var))) +
  labs(y='Score', x='Feature', size='Number of available features', 
       title='KBest Scores') +
  coord_flip()

```

As mentioned before 'loan_advances' just have 3 data points. It was therefore removed before a SelectKBest Analysis.

A univariate feature exploration was then performed with SelectKBest. As it can be seen from the figure the new created feature
'total_financial_benefits' has the highest score. The new feature 'fraction_of_messages_to_poi'
seems also promising, whereas the 'fraction_of_messages_to_poi' has a low score.

'to_messages' and 'from_messages' have a low score and are also used for the calculation of the new features, so they can be removed.

As the total_financial_benefits is the sum of 'salary', 'bonus', 'total_stock_value', 'exercised_stock_options'
these features are removed to avoid colinearity.



## Modeling

The selected features were applied to

```{r warning=FALSE, echo=FALSE, fig.height=10, fig.width=20, message=FALSE}
metrics_df %>% 
  gather(-classifier, -datatype, -tag, key = "var", value = "val")  %>%
  ggplot(aes(x = var, y=val)) +
    geom_point() +
    facet_wrap(~classifier, scales = "free", nrow = 2) + coord_flip()
```


### Tuning 

### Conclusion





