---
title: "R Notebook"
output: html_notebook
---

```{r}
library(ggplot2)
library(dplyr)
library(gridExtra)
```


```{r}
data("diamonds")
```

Let's start by examining two variables in the data set.
The scatterplot is a powerful tool to help you understand
the relationship between two continuous variables.

We can quickly see if the relationship is linear or not.
In this case, we can use a variety of diamond
characteristics to help us figure out whether
the price advertised for any given diamond is 
reasonable or a rip-off.

Let's consider the price of a diamond and it's carat weight.
Create a scatterplot of price (y) vs carat weight (x).

Limit the x-axis and y-axis to omit the top 1% of values.



```{r}

ggplot(data=diamonds, mapping=aes(y=price,x=carat)) +
  geom_jitter(alpha=1/40) +
  xlim(quantile(diamonds$carat,probs=c(0.00, .99)))+
  ylim(quantile(diamonds$price,probs=c(0.00, .99)))+
  geom_smooth(method='lm')
```

Create two histograms of the price variable
and place them side by side on one output image.

We’ve put some code below to get you started.

The first plot should be a histogram of price
and the second plot should transform
the price variable using log10.

Set appropriate bin widths for each plot.
ggtitle() will add a title to each histogram.

You can self-assess your work with the plots
in the solution video.


```{r}
p1 <- ggplot(data=diamonds, mapping = aes(x=price)) +
  geom_histogram(binwidth = 80)




p2 <- ggplot(data=diamonds, mapping = aes(x=price)) +
  geom_histogram(binwidth = 0.01) + scale_x_log10()

grid.arrange(p1,p2, nrow=1)
```

```{r cuberoot transformation}

cuberoot_trans = function(){
  trans_new('cuberoot',
            transform = function(x) {x^(1/3)},
            inverse = function(x){ x^3})
  }
```

Add a layer to adjust the features of the
scatterplot. Set the transparency to one half,
the size to three-fourths, and jitter the points.

If you need hints, see the Instructor Notes.
There are three hints so scroll down slowly if
you don’t want all the hints at once.

```{r}

ggplot(aes(carat, price), data = diamonds) + 
  geom_point(aes(color = clarity), alpha = 1/2, size = 3/4, position = 'jitter') + 
  scale_x_continuous(trans = cuberoot_trans(), limits = c(0.2, 3),
                     breaks = c(0.2, 0.5, 1, 2, 3)) + 
  scale_y_continuous(trans = log10_trans(), limits = c(350, 15000),
                     breaks = c(350, 1000, 5000, 10000, 15000)) +
  ggtitle('Price (log10) by Cube-Root of Carat')
```
```{r}

ggplot(aes(carat, price), data = diamonds) + 
  geom_point(alpha = 1/2, size = 3/4, position = 'jitter') + 
  scale_x_continuous(trans = cuberoot_trans(), limits = c(0.2, 3),
                     breaks = c(0.2, 0.5, 1, 2, 3)) + 
  scale_y_continuous(trans = log10_trans(), limits = c(350, 15000),
                     breaks = c(350, 1000, 5000, 10000, 15000)) +
  ggtitle('Price (log10) by Cube-Root of Carat')
```


```{r}

ggplot(aes(x = carat, y = price, color = cut), data = diamonds) + 
  geom_point(alpha = 0.5, size = 1, position = 'jitter') +
  scale_color_brewer(type = 'div',
                     guide = guide_legend(title = 'Clarity', reverse = T,
                                          override.aes = list(alpha = 1, size = 2))) +  
  scale_x_continuous(trans = cuberoot_trans(), limits = c(0.2, 3),
                     breaks = c(0.2, 0.5, 1, 2, 3)) + 
  scale_y_continuous(trans = log10_trans(), limits = c(350, 15000),
                     breaks = c(350, 1000, 5000, 10000, 15000)) +
  ggtitle('Price (log10) by Cube-Root of Carat and Clarity')
```


```{r A Bigger, Better Data Set}
library('bitops')
load('BigDiamonds.Rda')
```


Your task is to build five linear models like Solomon
did for the diamonds data set only this
time you'll use a sample of diamonds from the
diamondsbig data set.

Be sure to make use of the same variables
(logprice, carat, etc.) and model
names (m1, m2, m3, m4, m5).

To get the diamondsbig data into RStudio
on your machine, copy, paste, and run the
code in the Instructor Notes. There's
598,024 diamonds in this data set!

Since the data set is so large,
you are going to use a sample of the
data set to compute the models. You can use
the entire data set on your machine which
will produce slightly different coefficients
and statistics for the models.

This exercise WILL BE automatically graded.

You can leave off the code to load in the data.
We've sampled the data for you.
You also don't need code to create the table output of the models.
We'll do that for you and check your model summaries (R^2 values, AIC, etc.)

Your task is to write the code to create the models.
```{r}
library(memisc)
m1 <- lm(I(log(price)) ~ I(carat^(1/3)), data = diamondsbig)
m2 <- update(m1, ~ . + carat)
m3 <- update(m2, ~ . + cut)
m4 <- update(m3, ~ . + color)
m5 <- update(m4, ~ . + clarity)
mtable(m1, m2, m3, m4, m5)
```


```{r}

m6 <- lm(formula = I(log(price)) ~ I(carat^(1/3)) + carat + cut + color + 
    clarity, data = diamondsbig)
mtable(m6)
```

```{r}
thisDiamond = data.frame(carat = 1.00, cut = "V.Good",
                         color = "I", clarity="VS1")
modelEstimate = predict(m6, newdata = thisDiamond,
                        interval="prediction", level = .95)
```

