---
title: "problemset5"
output:
  pdf_document: default
  html_document: default
---
```{r}
library(ggplot2)
library(dplyr)
data(diamonds)
```


Create a histogram of diamond prices.
Facet the histogram by diamond color
and use cut to color the histogram bars.

The plot should look something like this.
http://i.imgur.com/b5xyrOu.jpg

Note: In the link, a color palette of type
'qual' was used to color the histogram using
scale_fill_brewer(type = 'qual')

This assignment is not graded and
will be marked as correct when you submit.

```{r}

ggplot(data=diamonds, mapping=aes(x=price, fill=cut))+
  geom_histogram() +
  scale_x_log10() +
  facet_wrap(~color) 
  

```


Create a scatterplot of diamond price vs.
table and color the points by the cut of
the diamond.

The plot should look something like this.
http://i.imgur.com/rQF9jQr.jpg

Note: In the link, a color palette of type
'qual' was used to color the scatterplot using
scale_color_brewer(type = 'qual')

This assignment is not graded and
will be marked as correct when you submit.

```{r}
ggplot(data=diamonds, mapping= aes(x=table, y=price, color=cut))+
  geom_point(alpha=1/5) + scale_x_continuous(limits = c(50,80))
```
```{r}
diamonds %>% group_by(cut) %>% summarise(low=quantile(table, probs=c(0.2)), high=quantile(table, probs=c(0.8)))
```

```{r}

summary(diamonds[diamonds$cut == 'Ideal',]$table)
summary(diamonds[diamonds$cut == 'Premium',]$table)
```
Create a scatterplot of diamond price vs.
volume (x * y * z) and color the points by
the clarity of diamonds. Use scale on the y-axis
to take the log10 of price. You should also
omit the top 1% of diamond volumes from the plot.

Note: Volume is a very rough approximation of
a diamond's actual volume.

The plot should look something like this.
http://i.imgur.com/excUpea.jpg

Note: In the link, a color palette of type
'div' was used to color the scatterplot using
scale_color_brewer(type = 'div')

This assignment is not graded and
will be marked as correct when you submit.

```{r}
diamonds$volume= diamonds$x * diamonds$y * diamonds$z
ggplot(data=subset(diamonds, volume < 0.99 * max(volume)), mapping=aes(y=price, x=volume, color=clarity))+
  geom_point(alpha=1/40) + scale_y_log10()+ scale_x_continuous(limit=c(0,450))
```

```{r}

ggplot(aes(x = volume, y = price), data = diamonds) + 
  geom_point(aes(color = clarity), alpha = 1/20) +
    scale_y_log10()+
  xlim(0, quantile(diamonds$volume, 0.99)) 
```


Many interesting variables are derived from two or more others.
For example, we might wonder how much of a person's network on
a service like Facebook the user actively initiated. Two users
with the same degree (or number of friends) might be very
different if one initiated most of those connections on the
service, while the other initiated very few. So it could be
useful to consider this proportion of existing friendships that
the user initiated. This might be a good predictor of how active
a user is compared with their peers, or other traits, such as
personality (i.e., is this person an extrovert?).

Your task is to create a new variable called 'prop_initiated'
in the Pseudo-Facebook data set. The variable should contain
the proportion of friendships that the user initiated.


```{r}

pf <- read.csv('../lesson4/data/pseudo_facebook.tsv', sep = '\t')
pf <- transform(pf, prop_initiated  = friendships_initiated/friend_count)
pf$year_joined <- floor(2014 - pf$tenure/365)
pf$year_joined.bucket <- cut(pf$year_joined, breaks = c(2004,2009,2011,2012,2014))
```
Create a line graph of the median proportion of
friendships initiated ('prop_initiated') vs.
tenure and color the line segment by
year_joined.bucket.

Recall, we created year_joined.bucket in Lesson 5
by first creating year_joined from the variable tenure.
Then, we used the cut function on year_joined to create
four bins or cohorts of users.

(2004, 2009]
(2009, 2011]
(2011, 2012]
(2012, 2014]

The plot should look something like this.
http://i.imgur.com/vNjPtDh.jpg
OR this
http://i.imgur.com/IBN1ufQ.jpg

This assignment is not graded and
will be marked as correct when you submit.
  geom_line(aes(color = year_joined.bucket), stat = 'summary', fun.y = median)
```{r}
ggplot(data=pf, mapping=aes(x=tenure, y=prop_initiated)) +
  geom_line(aes(color=year_joined.bucket), stats='summary', fun.y=mean)
```

```{r}

pf$year_joined <- floor(2014 - pf$tenure/365)
pf$year_joined.bucket <- cut(pf$year_joined, breaks = c(2004,2009,2011,2012,2014))
ggplot(aes(x = tenure, y = prop_initiated), data = pf) +
  geom_line(aes(color = year_joined.bucket), stat = 'summary', fun.y = mean)
```


