---
title: "Redwine quality analysis by Jens Laufer 2017/06/15"
output:
  html_document: default
  html_notebook: default
  pdf_document: default
---

## About

There are lot of myths about the quality of wine and what influences it. 

Within this data analysis physicochemical variables which are having an influence on the quality of red wine are examined. The quality of wine was rated by at least 3 wine experts. 

The dataset used within these data set is from can be [downloaded here](https://www.kaggle.com/piyushgoyal443/red-wine-dataset/downloads/red-wine-dataset.zip). 

The dataset contains 1599 red wines and 13 variables.



```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
library(ggplot2)
library(dplyr)
library(tidyr)
library(gridExtra)
library(GGally)
```

```{r echo=FALSE, Load_the_Data}
df = read.csv('./data/wineQualityReds.csv')
```



```{r echo=FALSE, message=FALSE, warning=FALSE, data_structure}
str(df)
```


### Description of attributes:


| Variable | Description | Unit | Data Type | Scale | Type |
| -------- | ----------- | ---- | --------- | ----- | ---- |
| X | Index of observation | | Unique Index | | |
| fixed acidity | lost acids involved with wine or fixed or nonvolatile (do not evaporate readily)| $g / l$ | Quantitative | Ratio | Input |
| volatile acidity | the amount of acetic acid in wine, which at too high of levels can lead to an unpleasant, vinegar taste | $g / l$ | Quantitative | Ratio | Input |
| citric acid | found in small quantities, citric acid can add 'freshness' and flavor to wines | $g / l$ | Quantitative | Ratio | Input |
| residual sugar | the amount of sugar remaining after fermentation stops, it's rare to find wines with less than 1 gram/liter and wines with greater than 45 grams/liter are considered sweet | $g / l$ | Quantitative | Ratio | Input |
| chlorides (sodium chloride) | the amount of salt in the wine | $g / l$ | Quantitative | Ratio | Input |
| free sulfur dioxide | the free form of SO2 exists in equilibrium between molecular SO2 (as a dissolved gas) and bisulfite ion; it prevents microbial growth and the oxidation of wine | $mg / l$ | Quantitative | Ratio | Input |
| total sulfur dioxide | amount of free and bound forms of S02; in low concentrations, SO2 is mostly undetectable in wine, but at free SO2 concentrations over 50 ppm, SO2 becomes evident in the nose and taste of wine | $mg / l$ | Quantitative | Ratio | Input |
| density | the density of water is close to that of water depending on the percent alcohol and sugar content | $g / cm^3$ | Quantitative | Ratio | Input |
| pH | describes how acidic or basic a wine is on a scale from 0 (very acidic) to 14 (very basic); most wines are between 3-4 on the pH scale | | Quantitative | Interval | Input |
| sulphates (potassium sulphate) | a wine additive which can contribute to sulfur dioxide gas (S02) levels, wich acts as an antimicrobial and antioxidant | $g / cm^3$ | Quantitative | Ratio | Input |
| alcohol | the percent alcohol content of the wine | $Vol. \%$ | Quantitative |  Ratio | Input |
| quality | quality rating from at least 3 wine experts | score between 0 and 10 | Qualitative | Ordinal | Output |



### Sample Data

```{r echo=FALSE, message=FALSE, warning=FALSE, example}
head(df)
```



# Univariate Plots Section


To get a first impression the distributions of all variables are plotted. As they are all quantitative data, this is done with histograms. The 'quality' will later transformed into a ordinal categorical variable.


```{r echo=FALSE, warning=FALSE, message=FALSE, fig.height=10, fig.width=20, warning=FALSE, Multivariate_Plots}
  df %>%
  gather(-X, key = "var", value = "value")  %>%
  ggplot(aes(x = value)) +
    geom_histogram() +
    facet_wrap(~var, scales = "free") 
```
 
 - ** Normally distributed **: PH and density  
 - ** Long tailed **: Chlorides, Residual Sugar
 - ** Positively Skewed **: alcohol, fixed.acidity, free.sulfur.dioxide, sulphate, total.sulfur.dioxide and volatile.acidity 
 

citric.acid needs to be analysed more as it has a lot peaks. Especially there many 0 values.

Based on [https://grapesandwine.cals.cornell.edu/sites/grapesandwine.cals.cornell.edu/files/shared/documents/Research-Focus-2011-3.pdf
](Sulphur Dioxide Content of Wines: the Role of Winemaking
and Carbonyl Compounds) there is maximum of sulphur dioxide in red wines of 150 mg/l for wines with a residual sugar level < 5 g/l and 200 mg/l with for residual sugar level > 5 g/l.
We created therfore a new categorical variable (is.max.total.sulfur.dioxide.exceed) for wines that exceed this limit. We are using the limit as the wines are from Portugal.


Based on [http://waterhouse.ucdavis.edu/whats-in-wine/volatile-acidity](Volatile Acidity in Wine) the limit for volatile acidity is 1.2 g/l in the US (and the EU) for red wine. A wine is spoiled above this limit.

```{r echo=FALSE, warning=FALSE, message=FALSE, Data_Modification}

df$quality_factor <- factor(df$quality, levels=seq(1,10,1), ordered = TRUE)
df$quality_bucket <- cut(df$quality, c(0,4,6,10), labels=c('low', 'average', 'high'))
df$is.max.total.sulfur.dioxide.exceed <- ifelse((df$total.sulfur.dioxide > 150 &
                                                  df$residual.sugar < 5) |
                                                  (df$total.sulfur.dioxide
                                                  > 200 &
                                                  df$residual.sugar > 5) , TRUE, FALSE)
df$is.max.volatile.acidity <- ifelse((df$volatile.acidity > 1.2) , TRUE, FALSE)
```


```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=5, fig.width=10, Univariate_Quality1}
df %>%
  ggplot(mapping=aes(x=quality_factor)) +
    geom_bar()
```


```{r}
df %>%
  filter(is.max.total.sulfur.dioxide.exceed == TRUE) %>%
  ggplot(mapping=aes(x=quality_factor)) +
   geom_bar()
  
```
There 7 wines which exceeds the limit for total sulphur dioxide. It's is suprising that 2 of these wines are high quality wines. Either these wines are for export outsite the EU or the winemakers are adding the sulfur dioxide to preserve it.



```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=5, fig.width=10, Univariate_Quality2}
df %>%
  filter(is.max.volatile.acidity == TRUE) %>%
  ggplot(mapping=aes(x=quality_factor)) +
    geom_bar()
```

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=5, fig.width=10, Univariate_Quality3}
df %>%
  ggplot(mapping=aes(x=quality_bucket)) +
    geom_bar()
```

### Alcohol

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=5, fig.width=10}
df %>%
  ggplot(mapping=aes(x=alcohol)) +
    geom_histogram(binwidth = 0.5) 
```

I am transforming the alcohol distribution by a log10 transformation. 

### Chlorides

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=5, fig.width=10}

df %>% 
  filter(chlorides > quantile(chlorides, probs=c(0.03)), 
         chlorides < quantile(chlorides, probs=c(0.97))) %>%
  ggplot(mapping=aes(x=chlorides)) +
    geom_histogram(binwidth = 0.022) +
    scale_x_log10(breaks=c(0.05, 0.07, 0.08, 0.1, 0.15, 0.2))

```

# Citric Acid

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=5, fig.width=10}
df %>% 
  ggplot(mapping=aes(x=citric.acid)) +
    geom_histogram(binwidth = 0.01) 
```

It is interesting taht we see peaks at 0.25, 0.5. You can see that winemakers are adding it during fermantation.

### Fixed Acidity




```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=5, fig.width=10}
df %>%
  ggplot(mapping=aes(x=fixed.acidity)) +
    geom_histogram(binwidth = 0.016) +
    scale_x_log10(breaks=seq(1,18,1)) 
```

### Free Sulfur Dioxide

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=5, fig.width=10}
df %>%
  ggplot(mapping=aes(x=free.sulfur.dioxide)) +
    geom_histogram(binwidth = 0.4) +
    scale_x_sqrt(breaks=c(1,2,5,10,15,20,25,30,40,50,70))
```
```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=5, fig.width=10}
df %>%
  filter(residual.sugar > quantile(residual.sugar, probs=c(0.025)), 
         residual.sugar < quantile(residual.sugar, probs=c(0.975))) %>%
  ggplot(mapping=aes(x=residual.sugar)) +
    geom_histogram(binwidth = 0.021) +
    scale_x_log10(breaks=seq(1,10,1))
```

### Sulphates


```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=5, fig.width=10}
df %>% 
  filter(sulphates > quantile(sulphates, probs=c(0.02)), 
         sulphates < quantile(sulphates, probs=c(0.98))) %>%
  ggplot(mapping=aes(x=sulphates)) +
    geom_histogram(binwidth = 0.021) +
    scale_x_log10(breaks=seq(0,2,0.1))
```

### Total Sulfur Dioxide

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=5, fig.width=10}
df %>% 
  filter(total.sulfur.dioxide > quantile(total.sulfur.dioxide, probs=c(0.02)), 
         total.sulfur.dioxide < quantile(total.sulfur.dioxide, probs=c(0.98))) %>%
  ggplot(mapping=aes(x=total.sulfur.dioxide)) +
    geom_histogram(binwidth = 0.09) +
    scale_x_log10(breaks=c(10,20,30,40,50,60,70,80,90,100,110,120,150))
```

### Volatile Acidity

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=5, fig.width=10}
df %>% 
  ggplot(mapping=aes(x=volatile.acidity)) +
    geom_histogram(binwidth = 0.033) +
    scale_x_log10(breaks=c(0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 1, 2))
```





> **Tip**: Make sure that you leave a blank line between the start / end of
each code block and the end / start of your Markdown text so that it is
formatted nicely in the knitted text. Note as well that text on consecutive
lines is treated as a single space. Make sure you have a blank line between
your paragraphs so that they too are formatted for easy readability.

# Univariate Analysis

> **Tip**: Now that you've completed your univariate explorations, it's time to
reflect on and summarize what you've found. Use the questions below to help you
gather your observations and add your own if you have other thoughts!

### What is the structure of your dataset?

### What is/are the main feature(s) of interest in your dataset?

### What other features in the dataset do you think will help support your \
investigation into your feature(s) of interest?

### Did you create any new variables from existing variables in the dataset?

### Of the features you investigated, were there any unusual distributions? \
Did you perform any operations on the data to tidy, adjust, or change the form \
of the data? If so, why did you do this?


# Bivariate Plots Section

> **Tip**: Based on what you saw in the univariate plots, what relationships
between variables might be interesting to look at in this section? Don't limit
yourself to relationships between a main output feature and one of the
supporting variables. Try to look at relationships between supporting variables
as well.



```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=10, fig.width=20, Bivariate_Plots_Correlation}
df[, -which(names(df) %in% c('X'))] %>%
   ggcorr(palette = "RdYlGn", label=TRUE)
```
alcohol, sulphates, volatile.acidy



```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=10, fig.width=20, Bivariate_Plots_scatter}
df %>%
  gather(-X, -quality_factor, -quality_bucket, -quality, key = "var", value = "value")  %>%
  ggplot(aes(x = quality, y = value)) +
    geom_point(alpha=0.3, position = 'jitter', color='orange') +
    geom_smooth() +
    facet_wrap(~var, scales = "free") 
```



```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=10, fig.width=20, Bivariate_Plots}
df %>%
  gather(-X, -quality_factor, -quality_bucket, -quality, key = "var", value = "value")  %>%
  ggplot(aes(x = quality_bucket, y=value)) +
    geom_point(alpha=0.3, position = 'jitter', color='orange') +
    geom_boxplot(alpha=0.5) +
    facet_wrap(~var, scales = "free") 
```

```{r, fig.height=30, fig.width=20}
ggpairs(df, columns = c('citric.acid', 'fixed.acidity', 'volatile.acidity', 'pH'))

```

```{r}

df[,+which(names(df) %in% c('sulphates', 'total.sulfur.dioxide', 'free.sulfur.dioxide'))] %>%
  ggpairs()

```




# Bivariate Analysis

> **Tip**: As before, summarize what you found in your bivariate explorations
here. Use the questions below to guide your discussion.

### Talk about some of the relationships you observed in this part of the \
investigation. How did the feature(s) of interest vary with other features in \
the dataset?

### Did you observe any interesting relationships between the other features \
(not the main feature(s) of interest)?

### What was the strongest relationship you found?


# Multivariate Plots Section

> **Tip**: Now it's time to put everything together. Based on what you found in
the bivariate plots section, create a few multivariate plots to investigate
more complex interactions between variables. Make sure that the plots that you
create here are justified by the plots you explored in the previous section. If
you plan on creating any mathematical models, this is the section where you
will do that.

```{r echo=FALSE, fig.height=10, fig.width=20}

df %>%
  ggplot(mapping=aes(x=alcohol, y=volatile.acidity, group=quality_bucket)) +
  geom_point(alpha=0.8, mapping=aes(fill=quality_factor), size=5, shape=21) +
  scale_fill_brewer(palette ='BuGn') +
  geom_boxplot(alpha=0.3) +
  facet_grid(~quality_bucket)

```

```{r echo=FALSE, fig.height=10, fig.width=20}

df %>%
  ggplot(mapping=aes(x=sulphates, y=alcohol)) +
  geom_point(alpha=0.7, mapping=aes(fill=quality_factor), size=5, shape=21) +
  scale_fill_brewer(palette ='RdYlGn') +
  facet_grid(~quality_bucket)

```

# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. Were there features that strengthened each other in terms of \
looking at your feature(s) of interest?

### Were there any interesting or surprising interactions between features?

### OPTIONAL: Did you create any models with your dataset? Discuss the \
strengths and limitations of your model.

------

# Final Plots and Summary

> **Tip**: You've done a lot of exploration and have built up an understanding
of the structure of and relationships between the variables in your dataset.
Here, you will select three plots from all of your previous exploration to
present here as a summary of some of your most interesting findings. Make sure
that you have refined your selected plots for good titling, axis labels (with
units), and good aesthetic choices (e.g. color, transparency). After each plot,
make sure you justify why you chose each plot by describing what it shows.

### Plot One
```{r echo=FALSE, Plot_One}

```

### Description One


### Plot Two
```{r echo=FALSE, Plot_Two}

```

### Description Two


### Plot Three
```{r echo=FALSE, Plot_Three}

```

### Description Three

------

# Reflection

> **Tip**: Here's the final step! Reflect on the exploration you performed and
the insights you found. What were some of the struggles that you went through?
What went well? What was surprising? Make sure you include an insight into
future work that could be done with the dataset.

> **Tip**: Don't forget to remove this, and the other **Tip** sections before
saving your final work and knitting the final report!
