---
title: "Redwine quality analysis by Jens Laufer 2017/06/15"
output:
  pdf_document: default
  html_notebook: default
  html_document: default
---

## About

There are lot of myths about the quality of wine and what influences it. 

Within this data analysis physicochemical properties which are having an influence on the quality of red wine are examined. 

The dataset used within these data analysis is from [[Cortez et al., 2009]](http://www3.dsi.uminho.pt/pcortez/wine/) and can be downloaded here [winequality.zip](http://www3.dsi.uminho.pt/pcortez/wine/winequality.zip). 

The dataset contains 1599 red wines and 13 variables from Portugal. The quality was rated by at least 3 wine experts. 



```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
library(ggplot2)
library(dplyr)
library(tidyr)
library(gridExtra)
library(GGally)
library(knitr)
```

```{r echo=FALSE, Load_the_Data}
df = read.csv('./data/wineQualityReds.csv')
```



```{r echo=FALSE, message=FALSE, warning=FALSE, data_structure}
str(df)
```


### Description of attributes:


| Variable | Description | Unit | Data Type | Scale | Type |
| -------- | ----------- | ---- | --------- | ----- | ---- |
| X | Index of observation | | Unique Index | | |
| fixed acidity | lost acids involved with wine or fixed or nonvolatile (do not evaporate readily)| $g / l$ | Quantitative | Ratio | Input |
| volatile acidity | the amount of acetic acid in wine, which at too high of levels can lead to an unpleasant, vinegar taste | $g / l$ | Quantitative | Ratio | Input |
| citric acid | found in small quantities, citric acid can add 'freshness' and flavor to wines | $g / l$ | Quantitative | Ratio | Input |
| residual sugar | the amount of sugar remaining after fermentation stops, it's rare to find wines with less than 1 gram/liter and wines with greater than 45 grams/liter are considered sweet | $g / l$ | Quantitative | Ratio | Input |
| chlorides (sodium chloride) | the amount of salt in the wine | $g / l$ | Quantitative | Ratio | Input |
| free sulfur dioxide | the free form of SO2 exists in equilibrium between molecular SO2 (as a dissolved gas) and bisulfite ion; it prevents microbial growth and the oxidation of wine | $mg / l$ | Quantitative | Ratio | Input |
| total sulfur dioxide | amount of free and bound forms of S02; in low concentrations, SO2 is mostly undetectable in wine, but at free SO2 concentrations over 50 ppm, SO2 becomes evident in the nose and taste of wine | $mg / l$ | Quantitative | Ratio | Input |
| density | the density of water is close to that of water depending on the percent alcohol and sugar content | $g / cm^3$ | Quantitative | Ratio | Input |
| pH | describes how acidic or basic a wine is on a scale from 0 (very acidic) to 14 (very basic); most wines are between 3-4 on the pH scale | | Quantitative | Interval | Input |
| sulphates (potassium sulphate) | a wine additive which can contribute to sulfur dioxide gas (S02) levels, which acts as an antimicrobial and antioxidant | $g / cm^3$ | Quantitative | Ratio | Input |
| alcohol | the percent alcohol content of the wine | $Vol. \%$ | Quantitative |  Ratio | Input |
| quality | quality rating from at least 3 wine experts | score between 0 and 10 | Qualitative | Ordinal | Output |



### Sample Data Entries

```{r echo=FALSE, message=FALSE, warning=FALSE, example}
kable(head(df))
```



# Univariate Plots Section


To get a first impression the distributions of all variables are plotted. As they are all quantitative data, this is done with histograms. The 'quality' will later transformed into a ordinal categorical variable.


```{r echo=FALSE, warning=FALSE, message=FALSE, fig.height=10, fig.width=20, warning=FALSE, Multivariate_Plots}
  df %>%
  gather(-X, key = "var", value = "value")  %>%
  ggplot(aes(x = value)) +
    geom_histogram() +
    facet_wrap(~var, scales = "free") 
```

__Normally distributed__: PH and density  

__Long tailed__: Chlorides, Residual Sugar

__Positively Skewed__: alcohol, fixed.acidity, free.sulfur.dioxide, sulphate, total.sulfur.dioxide and volatile.acidity 
 
__Uniform__: citric.acid


Based on [[Sulphur Dioxide Content of Wines: the Role of Winemaking
and Carbonyl Compounds]](https://grapesandwine.cals.cornell.edu/sites/grapesandwine.cals.cornell.edu/files/shared/documents/Research-Focus-2011-3.pdf
) there is a limit for sulphur dioxide in red wines in the European Union of 150 mg/l for wines with a residual sugar level < 5 g/l and 200 mg/l for residual sugar level > 5 g/l.

A new categorical variable (is.max.total.sulfur.dioxide.exceed) for wines that exceed this limit is therefore created. The limits from the EU are used as the wines are from Portugal.


Based on [[Volatile Acidity in Wine]](http://waterhouse.ucdavis.edu/whats-in-wine/volatile-acidity) the limit for volatile acidity is 1.2 g/l in the US (and the EU) for red wine. A wine is spoiled above this limit. A new categorical (is.max.volatile.acidity) is created.

Based on [Wine Sweetness](https://en.wikipedia.org/wiki/Sweetness_of_wine#Residual_sugar) a new categorical variable for wine sweetness is also created. 



```{r echo=FALSE, warning=FALSE, message=FALSE, Data_Modification}

df$quality_factor <- factor(df$quality, levels=seq(1,10,1), ordered = TRUE)
df$quality_bucket <- cut(df$quality, c(0,4,6,10), labels=c('low', 'average', 'high'))
df$is.max.total.sulfur.dioxide.exceed <- ifelse((df$total.sulfur.dioxide > 150 &
                                                  df$residual.sugar < 5) |
                                                  (df$total.sulfur.dioxide
                                                  > 200 &
                                                  df$residual.sugar > 5) , TRUE, FALSE)
df$is.max.volatile.acidity <- ifelse((df$volatile.acidity > 1.2) , TRUE, FALSE)

df$sweetness = cut(df$residual.sugar,
                   c(0,4,12,45,1000),
                   labels = c('dry',
                              'medium dry', 'medium', 'sweet'))
```


### Quality

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=5, fig.width=10, Univariate_Quality1}
df %>%
  ggplot(mapping=aes(x=quality_factor)) +
    geom_bar()
```

```{r}
summary(df$quality)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=5, fig.width=10, Univariate_Quality3}
df %>%
  ggplot(mapping=aes(x=quality_bucket)) +
    geom_bar()
```
```{r}
table(df$quality_bucket)
```


Most of the wines are average. There are some high quality and low quality wines.

There are no wines with quality below 3 and no wines above 8. Most wine fall into the quality factor 5-6. The median of the quality is 6.




### Sweetness

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=5, fig.width=10, Univariate_Quality4}
df %>%
  ggplot(mapping=aes(x=sweetness)) +
    geom_bar()
```

```{r}
table(df$sweetness)
```

Most of the wines are dry and medium dry. There are some medium wines and no sweet wine in the dataset.



### Alcohol

Alcohol in wine is created during fermantation of the grapes.


```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=5, fig.width=10}
df %>%
  ggplot(mapping=aes(x=alcohol)) +
    geom_histogram(binwidth = 0.5) 
```


A log10 transformation is done on the alcohol distribution to get a normal-like distribution.


```{r}
summary(df$alcohol)
```

The median alcohol is 10.2 Vol. %. Most wines are having between 9.5 and 11.1 Vol. % of alcohol.


### Chlorides

Chlorides in wine is actually how much salt a wine has. It is influenced by terrior [[Coli et al]](http://www.scielo.br/scielo.php?script=sci_arttext&pid=S0101-20612015000100095)

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=5, fig.width=10}

df %>% 
  filter(chlorides > quantile(chlorides, probs=c(0.03)), 
         chlorides < quantile(chlorides, probs=c(0.97))) %>%
  ggplot(mapping=aes(x=chlorides)) +
    geom_histogram(binwidth = 0.022) +
    scale_x_log10(breaks=c(0.05, 0.07, 0.08, 0.1, 0.15, 0.2))

```

A log10-transformation was done on the chlorides distribution to get a almost normal distribution.

```{r}
summary(df$chlorides)
```

The most wines has a chlorides concentration between 0.07 and 0.09 g/l and a median of 0.079 g/l



### Citric Acid

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=5, fig.width=10}
df %>% 
  ggplot(mapping=aes(x=citric.acid)) +
    geom_histogram(binwidth = 0.005) +
    scale_x_continuous(breaks=seq(0,1,0.05))
```



```{r}
tail(sort(table(df$citric.acid)))
```


There are peaks at 0, 0.02, 0.24, 0.26, 0.49.


Naturally there is about 0.1 g/l or less citric acid in red wine. So the peaks of 0 and 0.02 are actually natural citric acidity. 

Winemakers add citric acid to give the wine more freshness or to stabilize the wine according to [[Wine Science]](https://books.google.de/books?id=9En27zEnVgwC&pg=PA367&lpg=PA367&dq=reduction+metal+wine+citric+acid&source=bl&ots=wplhimu9xq&sig=43hhIlDeYJ74-EROlBsOJQyRraU&hl=en&sa=X&ved=0ahUKEwjys-G__cnUAhXHY1AKHQxnBvMQ6AEITjAG#v=onepage&q=reduction%20metal%20wine%20citric%20acid&f=false). Citric acid limits the occurrence of ferric casse. They typically add between 0.5 and 0.8 g/l for stabilisation. The peak at 0.49 might due this addition.

On the other hand according  to [[Säuerung]](http://www.dlr.rlp.de/Internet/global/themen.nsf/0/5B71C70F91852E04C12576500027D0E9/$FILE/2009-10-09%20KIS%20%207.pdf) a citric acid concentration over 0.3-0.4 g/l has a bad influence on the taste. 



### Fixed Acidity

According to [[Acids in Wine]](https://en.wikipedia.org/wiki/Acids_in_wine) most of the acids in wine are fixed acids. In the grapes tartaric, malic and citric acid is found. During winemaking butyric, lactic and succinic acids are playing a role.


```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=5, fig.width=10}
df %>%
  ggplot(mapping=aes(x=fixed.acidity)) +
    geom_histogram(binwidth = 0.016) +
    scale_x_log10(breaks=seq(1,18,1)) 
```

```{r}
summary(df$fixed.acidity)
```


A log10 transformation was applied to get a almost normal distribution. Most of the wines are having a fixed acidity between 7.10 and 9.20 g/l and median of 7.90 g/l.


### Free Sulfur Dioxide


The free form of SO2 exists in equilibrium between molecular SO2 (as a dissolved gas) and bisulfite ion; it prevents microbial growth and the oxidation of wine

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=5, fig.width=10}
df %>%
  ggplot(mapping=aes(x=free.sulfur.dioxide)) +
    geom_histogram(binwidth = 0.4) +
    scale_x_sqrt(breaks=c(1,2,5,10,15,20,25,30,40,50,70))
```
```{r}
summary(df$free.sulfur.dioxide)
```

The concentration of free sulphure dioxid is between 7.00 and 21.00 mg/l with a median of 14.00 mg/l. 


### Residual Sugar

Residual sugar is the concentration of sugar, when fermantation stops. Winemakers sometimes cool down the wine to artificially stop the fermantation.



```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=5, fig.width=10}
df %>%
  filter(residual.sugar > quantile(residual.sugar, probs=c(0.025)), 
         residual.sugar < quantile(residual.sugar, probs=c(0.975))) %>%
  ggplot(mapping=aes(x=residual.sugar)) +
    geom_histogram(binwidth = 0.021) +
    scale_x_log10(breaks=seq(1,10,1))
```
```{r}
summary(df$residual.sugar)
```



A log10 transformation is done to get a normal distribution plot. Most wines are having a residual.sugar between 1.9 and 2.6 g/l with a median of 2.2 g/l.

There could not found signs in the distribution, if fermantation stopped naturally or was stopped.


### Sulphates

A wine additive which can contribute to sulfur dioxide gas (S02) levels, which acts as an antimicrobial and antioxidant

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=5, fig.width=10}
df %>% 
  filter(sulphates > quantile(sulphates, probs=c(0.02)), 
         sulphates < quantile(sulphates, probs=c(0.98))) %>%
  ggplot(mapping=aes(x=sulphates)) +
    geom_histogram(binwidth = 0.021) +
    scale_x_log10(breaks=seq(0,2,0.1))
```

```{r}
summary(df$sulphates)
```

50% of the wines are having a suplphates concentration between 0.55 and 0.73. The median is 0.62. A log10 transformation was applied to the plot.


### Total Sulfur Dioxide

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=5, fig.width=10}
df %>% 
  filter(total.sulfur.dioxide > quantile(total.sulfur.dioxide, probs=c(0.02)), 
         total.sulfur.dioxide < quantile(total.sulfur.dioxide, probs=c(0.98))) %>%
  ggplot(mapping=aes(x=total.sulfur.dioxide)) +
    geom_histogram(binwidth = 0.09) +
    scale_x_log10(breaks=c(10,20,30,40,50,60,70,80,90,100,110,120,150))
```

```{r}
summary(df$total.sulfur.dioxide)
```


A log10-transformation was applied to get a normal distribution. 50% of the wines are having a concentration between 22 and 62 mg/l. The median is 38 mg/l.


```{r}
df %>%
  filter(is.max.total.sulfur.dioxide.exceed == TRUE) %>%
  summarise(num.max.total.sulfur.dioxide.exceed = n())
  
```



There are 7 wines which exceeds the limit for total sulphur dioxide.



### Volatile Acidity

According to [Acids in wine](https://en.wikipedia.org/wiki/Acids_in_wine#Citric_acid) most of acids in wines are fixed acids with the exception of acetic acids, mostly found in vinegar, which is volatile and can contribute to the wine fault known. As mentioned before there are limits for the concentration of volatile acids in wine.


```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=5, fig.width=10}
df %>% 
  ggplot(mapping=aes(x=volatile.acidity)) +
    geom_histogram(binwidth = 0.033) +
    scale_x_log10(breaks=c(0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 1, 2))
```
```{r}
summary(df$volatile.acidity)
```


50% of the wines are having a concentration of volatile acidity between 0.39 and 0.64 with a median of 0.52 g/l.


```{r}
df %>%
  filter(is.max.volatile.acidity == TRUE) %>%
  summarise(num.is.max.volatile.acidity = n())
  
```


There are for 4 wines which exceeds the limit for volatile acidity. 


# Univariate Analysis

The dataset contains 1599 observations with 13 variables. 'quality' is the outcome variable, as the study is examining the influence of physicochemical properties on wines. 'X' is a unique key, so it can be skipped The other variables are income variables. Some new variables were created:

   - **quality_factor**: quality as categorical variable with a ordinale scale
   
   - **quality_bucket**: quality as categorical variable with levels 'low', 'average' and 'high'
   
   - **sweetness**: categorical variable for wine sweetness based on residual sugar levels
   
   - **is.max.total.sulfur.dioxide.exceed**: binary variable for wines with total sulfur dioxide levels over the limit
   
   - **is.max.volatile.acidity**: binary variable for wines wwith volatile acidity over the limit
   
Most of the variables were positively skewed or long tail distritutions, therefore log10 tranformations were mostly applied to get a kind of normal distribution. 

The distribution of citric acid was unusual with some spikes. The spikes are coming most probably because some wine makers are adding citric acid to the wine and the wine has as well natural ciric acidity.

There were 7 wines which are exceeding the limits for volatile acidity and 4 wines exceeding the limit for total sulfur exceeding. It will be interesting especially how these 11 wines are rated from the wine experts. 

Some of the variables are related like alcohol and residual sugar or pH, volatile and free acidity, citric acidity  or sulphates, total sulfur dioxide and free total sulfur dioxide. It's interested how these variables are correlated.

From the income variables some of the variables are suspected to have more influence on the wines than others: It's suspected that chlorides and sulphates are having an influences as the are salts, as well as all kind of acidity, the residual sugar and alcohol as all these "components" might be tasted.



# Bivariate Plots Section

> **Tip**: Based on what you saw in the univariate plots, what relationships
between variables might be interesting to look at in this section? Don't limit
yourself to relationships between a main output feature and one of the
supporting variables. Try to look at relationships between supporting variables
as well.



```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=10, fig.width=20, Bivariate_Plots_Correlation}
df[, -which(names(df) %in% c('X'))] %>%
   ggcorr(palette = "RdYlGn", label=TRUE)
```



```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=10, fig.width=20, Bivariate_Plots_Scatter}
df %>%
  gather(-X, -quality_factor, -quality_bucket, 
         -quality, -is.max.total.sulfur.dioxide.exceed,
         -is.max.volatile.acidity, -sweetness, key = "var", value = "value")  %>%
  ggplot(aes(x = quality_factor, y = value)) +
    geom_point(alpha=0.3, position = 'jitter', color='orange') +
    geom_boxplot(alpha=0.5) +
    facet_wrap(~var, scales = 'free') 
```



```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=10, fig.width=20, Bivariate_Plots}
df %>%
  gather(-X, -quality_factor, -quality_bucket, 
         -quality, -is.max.total.sulfur.dioxide.exceed,
         -is.max.volatile.acidity, -sweetness, key = "var", value = "value")  %>%
  ggplot(aes(x = quality_bucket, y=value)) +
    geom_point(alpha=0.3, position = 'jitter', color='orange') +
    geom_boxplot(alpha=0.5) +
    facet_wrap(~var, scales = "free") 
```


### Alcohol vs Quality

```{r, fig.height=10, fig.width=20}
df %>%
  ggplot(mapping=aes(x=quality, y=alcohol)) +
  geom_point(alpha=0.2, position = 'jitter', size=7) +
  geom_smooth(method = 'lm') +
  scale_x_continuous(breaks=seq(1,10,1))
```


### Volatile Acidity vs Quality

```{r, fig.height=10, fig.width=20}
df %>%
  ggplot(mapping=aes(x=quality, y=volatile.acidity)) +
  geom_point(alpha=0.2, position = 'jitter', size=7) +
  geom_smooth(method = 'lm') +
  scale_x_continuous(breaks=seq(1,10,1))
```


```{r}
df %>%
  filter(is.max.volatile.acidity == TRUE) %>%
  ggplot(mapping=aes(x=quality_bucket)) +
  geom_bar()
```



There are for 4 wines which exceeds the limit for volatile acidity.

These wines are having are rated with low and lower average. It seems that too much volatile acidity can be tasted. 


### Sulphates vs Quality

```{r, fig.height=10, fig.width=20, message=FALSE, warning=FALSE}
df %>%
  ggplot(mapping=aes(x=quality, y=sulphates)) +
  geom_point(alpha=0.2, position = 'jitter', size=4) +
  geom_smooth(method='lm') +
  scale_x_continuous(breaks=seq(0,10,1))
  
```





### citric.acidity vs quality 


```{r, fig.height=10, fig.width=20, message=FALSE, warning=FALSE}

df %>%
  ggplot(mapping=aes(x=quality, y=citric.acid)) +
  geom_point(alpha=0.2, position = 'jitter', size=6) +
  geom_smooth(method='lm') +
  scale_x_continuous(breaks=seq(0, 10, 1)) +
  scale_y_continuous(breaks=seq(0, 1, 0.1))
```




### total.sufur.dioxide vs quality

```{r, fig.height=10, fig.width=20, message=FALSE, warning=FALSE}
df %>%
  ggplot(mapping=aes(x=quality, y=total.sulfur.dioxide)) +
  geom_point(alpha=0.2, position = 'jitter', size=4) +
  geom_smooth(method='lm') +
  scale_x_continuous(breaks=seq(0,10,1))
  
```

```{r, fig.height=10, fig.width=20, message=FALSE, warning=FALSE}
df %>%
  filter(is.max.total.sulfur.dioxide.exceed == TRUE) %>%
  ggplot(mapping=aes(x=quality_bucket)) +
  geom_bar()
```


### density vs quality

```{r, fig.height=10, fig.width=20, message=FALSE, warning=FALSE}

```




# Bivariate Analysis

> **Tip**: As before, summarize what you found in your bivariate explorations
here. Use the questions below to guide your discussion.

### Talk about some of the relationships you observed in this part of the \
investigation. How did the feature(s) of interest vary with other features in \
the dataset?

### Did you observe any interesting relationships between the other features \
(not the main feature(s) of interest)?

### What was the strongest relationship you found?


# Multivariate Plots Section

> **Tip**: Now it's time to put everything together. Based on what you found in
the bivariate plots section, create a few multivariate plots to investigate
more complex interactions between variables. Make sure that the plots that you
create here are justified by the plots you explored in the previous section. If
you plan on creating any mathematical models, this is the section where you
will do that.

```{r echo=FALSE, fig.height=10, fig.width=20}

df %>%
  ggplot(mapping=aes(x=alcohol, y=volatile.acidity, group=quality_bucket)) +
  geom_point(alpha=0.8, mapping=aes(fill=quality_factor), size=5, shape=21) +
  scale_fill_brewer(palette ='BuGn') +
  geom_boxplot(alpha=0.3) +
  facet_grid(~quality_bucket)

```

```{r echo=FALSE, fig.height=10, fig.width=20}

df %>%
  ggplot(mapping=aes(x=sulphates, y=alcohol)) +
  geom_point(alpha=0.7, mapping=aes(fill=quality_factor), size=5, shape=21) +
  scale_fill_brewer(palette ='RdYlGn') +
  facet_grid(~quality_bucket)

```


```{r echo=FALSE, fig.height=10, fig.width=20}

df %>%
  ggplot(mapping=aes(x=quality_bucket, y=citric.acid)) +
  geom_point(alpha=0.4, position = 'jitter', mapping=aes(fill=quality_factor), size=5, shape=21) 
  scale_fill_brewer(palette ='RdYlGn') 

```




# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. Were there features that strengthened each other in terms of \
looking at your feature(s) of interest?

### Were there any interesting or surprising interactions between features?

### OPTIONAL: Did you create any models with your dataset? Discuss the \
strengths and limitations of your model.

------

# Final Plots and Summary

> **Tip**: You've done a lot of exploration and have built up an understanding
of the structure of and relationships between the variables in your dataset.
Here, you will select three plots from all of your previous exploration to
present here as a summary of some of your most interesting findings. Make sure
that you have refined your selected plots for good titling, axis labels (with
units), and good aesthetic choices (e.g. color, transparency). After each plot,
make sure you justify why you chose each plot by describing what it shows.

### Plot One
```{r echo=FALSE, Plot_One}

```

### Description One


### Plot Two
```{r echo=FALSE, Plot_Two}

```

### Description Two


### Plot Three
```{r echo=FALSE, Plot_Three}

```

### Description Three

------

# Reflection

> **Tip**: Here's the final step! Reflect on the exploration you performed and
the insights you found. What were some of the struggles that you went through?
What went well? What was surprising? Make sure you include an insight into
future work that could be done with the dataset.

> **Tip**: Don't forget to remove this, and the other **Tip** sections before
saving your final work and knitting the final report!
